### 2017-04-11 paper/samplebody-journals.tex
added tex doc and pdf

\section{Introduction/Motivation}

\quad Yelp is a website which allows users to create an account and review businesses. Our goal is to analyze data from Yelp to gain further insight for business owners to develop a satisfactory experience for their customers. Yelp has a public dataset available on their website with information about users, businesses, and reviews.

\quad Our  rst objective is to determine if certain attributes are indica- tive of business success. If we can find a way to predict whether or not a business will do well later on in its operation, that information will be useful to businesses looking to improve their model.

\quad Our second objective is to analyze the text of reviews for a sen- timent. The sentiment analysis will show the emotions conveyed by the reviewer as being pleased or angry at their experience with the business. Using the results of this analysis we hope to be able to determine if certain times of day, city locations, or other factors contribute to the experience of a customer. This information could also contribute to a business model. As an example, if certain loca- tions are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.

\quad Our last objective is to  nd recurring problems in businesses. We will analyze review text to  nd the speci c issues that people are having with businesses. If a business has many reviews which all indicate a similar issue, then that is something a business owner would want to be aware of so that they can eliminate the problem.

% Head 1
\section{Prior Work}

Since our problem involves predictions surrounding businesses, we sought out similar research to guide our investigation. Thus, we focus our literature review on studies that utilized some form of sentiment analysis, constructed predictive models, and/or conducted meta-analysis of trends in the aforementioned sentiment, rating, or business interest.

% Head 2
\subsection{Predicting Business Attention}

A previous winner, Hood et al., published a paper on assessing a business?s current rating/review state and future expectations with respect to review numbers on Yelp. The bulk of their work centered around the creation of additional features for the data (e.g. number of similar businesses within 1km, features of subsets of reviews such as average rating and number of unique users). From there they utilized various feature reduction and clustering techniques to build a prediction model. They were able to identify features that o ered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.

% Head 3
\subsection{Latent Subtopics}

Huang et al. (2013) utilized probabilistic models to discover under- lying subtopics within Yelp reviews. This o ered a framework to categorize the reviews based on review text. From here, they were able to analyze the prevalence of various star ratings within each category. In this way, they could advise business owners as to the common trends among good and bad reviews that may not be ap- parent without considering the review text. McAuley and Leskovec built on this idea by combining such latent review topics with hidden rating driving factors, allowing them construct a rating prediction model.

\subsection{Reviews and Revenue}

In one of the more in uential papers on the subject, Michael Luca showed a correlation between Yelp rating and revenue, highlighting the fact that rounding to half stars could signi cantly impact two similarly-rated restaurants if their averages fell on either side of the divide (e.g. 3.24 and 3.26). He goes on to build a model of market response based on review volume as well as the impact of user expertise. He concluded that such responses are consistent with a Bayesian learning model.

\section{Proposed Work and Dataset Information}

\subsection{Proposed Work}

The dataset has many optional attributes. There is an attribute called 'attributes' that is a list of optional attributes. The dataset is inconsistent and cannot be relied upon for each business. For instance, one of these attributes is ?Bike Parking?. This attribute is either true, false, or nonexistent. Speci cally the fact that this attribute might not be available for multiple businesses makes it unreliable.

\quad In addition, the data is not only restaurants but any kind of busi- ness. As a result they have another piece of data which is the cat- egories list. For each business this attribute attempts to classify the business. However some categories are extremely broad and the same type of business can be described in di erent ways. One example is that there is a bakery, but it?s also categorized as food. Many businesses will sell food, but the category of food will not be in their list of categories and instead it will be  lled with more speci c categories like Italian food.

\quad The majority of the data cleaning process will be involved in handling missing data. It is possible to  ll in these missing values with default values like false but this can be dangerous to make assumptions about what businesses have and don?t have.

\quad The Yelp dataset contains up to one million di erent attributes about businesses from 11 di erent cities in 4 countries. The majority of these attributes will be unnecessary so there will be a large e ort in data reduction. For instance there are photographs included, but our questions does not require image analysis. There also attributes such as postal codes that are unnecessary because latitude and longitude is more accurate.

\subsection{Dataset}

The Yelp dataset comes from the Yelp Dataset Challenge. This is an ongoing challenge and this is the ninth round. The challenge offers \$5,000 to the winner and the  final submission is an academic paper.

\quad We will be considering the following from the Yelp dataset:

\begin{itemize}
	\item{4.1 million reviews}
	\begin{verbatim}
	"review_id":(encrypted review id),
	"user_id":(encrypted user id),
	"business_id":(encrypted business id)"
	"stars":(star rating, rounded to half-stars),
	"date":(date, formatted: YYYY-MM-DD),
	"text":"review text",
	"useful":('useful' vote count),
	"funny":('funny' vote count),
	"cool": ('cool' vote count),
	"type": "review"
	\end{verbatim}
	\item{Informationon144,000businessesin11citiesfrom4coun- tries (incl. the US, Germany, Canada, and the U.K. (with 1.1 million business attributes)}
	\begin{verbatim}
	"business_id":(encrypted business id),
	"name":"business name",
	"neighborhood": "hood name",
	"address":"full address",
	"city":"city",
	"state":"state"(if applicable),
	"postal code":"postal code",
	"latitude":latitude,
	"longitude":longitude,
	"stars":(star rating, rounded to half-stars),
	"review_count":(number of reviews),
	"is_open":0/1 (closed/open),
	"attributes":[(localized attribute tags)],
	"categories":[(localized category names)],
	"hours":[(hours strings)],
	"type": "business"
	\end{verbatim}
	\item{Over 1M users}
	\begin{verbatim}
	"user_id":(encrypted user id),
	"name":"first name",
	"review_count":(number of reviews),
	"yelping_since": (date, formatted: YYYY-MM-DD),
	"friends":[(encrypted user ids)],
	"useful":('useful' vote count sent by user),
	"funny":('funny' vote count sent by user),
	"cool": ('cool' vote count sent by user),
	"fans":"number of fans the user has",
	"elite":["an array of years the user was elite"],
	"average_stars":floating point average like 4.31,
	"compliment_type": (compliment count),
		...(same for each compliment type)...
	"type":"user"
	\end{verbatim}
	\item{Over 200,000 user-submitted photographs, 947,000 tips, and aggregated check-in data}
\end{itemize}

\section{Methods and Tools}

\quad The tools and methods used for this project will include a database for storing and integrating the data, libraries and modules for cleaning and preprocessing, and various tools for transformation, mining, pattern evaluation, and knowledge presentation. Given the json-formatted data, we've opted for using a NoSQL database like MongoDB or Cassandra. Fortunately, there are a number of options, considering our educational status, that will allow large storage volume as well as low latency read/write access. Hosting in such a manner will requireA number of Python libraries like Dora and FTFY (?fixes text for you?) intended for data cleaning and preprocessing will be used for these purposes. Yelp has already performed a substantial amount of cleaning on the data, so it is likely that minimal additional cleaning will be necessary, save for replacing missing values as mentioned above.

\quad Numpy and Scipy will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation. Finding angry reviews and recurring problems will make heavy use of NLP, so the team will rely on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, and possibly some custom supervised text classi ers trained with the Facebook fastText pretrained word vectors to perform sentiment analysis on the review text.

\section{Evaluation}

\subsection{Methodology}

\quad Evaluating our predictions as to whether a business success will hinge on comparing the star rating (from one to  ve stars) and our predicted star rating based on the other attributes in the review, such as day of the week and month, season, or location. This could be done by training an ML algorithm like Kernel Ridge regression on 80\% of the data and testing it on the remaining 20\%, and then adjusting parameters to minimize the error between the actual star ratings and our predicted star ratings for the testing data.

\quad Evaluating the sentiment analysis of the review text will likely be simple given that we will be using tools like Weka and Microsoft Sentiment Analysis (see following section) to perform sentiment analysis rather than creating entirely custom algorithms. These tools incorporate helpful metrics to analyze the error of a particular output. Alternatively, we can manually verify the sentiment of a (relatively small) number of reviews by reading them and comparing our human judgements of the sentiment to the output of the tools. For evaluating recurring problems, we will employ a similar approach, except we will group reviews by business and analyze them as a time series to and recurring issues.

\quad For all three of the objective questions, employing association rule frameworks that use support-confidence frameworks will be helpful to and strong associations, while null-invariant measures like the Kulczynski measure or the cosine measure will be helpful to determine how interesting these associations are.

\section{Next Steps}

\quad Some major milestones before Spring Break (March 27th) will include organizing and integrating all the data in a relational database, as well as cleaning and preprocessing the data. After these steps have been completed, it will be much easier to begin preliminary investigations into our main questions for the project, such as reducing the entire set of attributes into just those that will be mined for each question. Milestones after Spring Break prior to April 11th will include using the tools listed above to answer two of our objective questions, while work after that will focus on answering the remaining question and writing the final report.

% Start of "Sample References" section

%\section{Typical references in new ACM Reference Format}
%A paginated journal article \cite{Abril07}, an enumerated
%journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
%a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
%\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
%followed by the same example, however we only output the series if the volume number is given
%\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
%a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
%in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
%an article in a proceedings (of a conference, symposium, workshop for example)
%(paginated proceedings article) \cite{Andler79}, a proceedings article
%with all possible elements \cite{Smith10}, an example of an enumerated
%proceedings article \cite{VanGundy07},
%an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
%a master's thesis: \cite{anisi03}, an online document / world wide web
%resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
%and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
%work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
%\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
%'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
%Boris / Barbara Beeton: multi-volume works as books
%\cite{MR781536} and \cite{MR781537}.
%
%A couple of citations with DOIs: \cite{2004:ITE:1009386.1010128,
%  Kirschmer:2010:AEI:1958016.1958018}.

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

### 2017-04-11 paper/samplebody-journals.tex
added images and results sections to paper

@@
 
+\section{Progress}
+
+\quad Our most significant hurdle was in getting our data into a tool for some initial experimentation.
+While JSON formatting is extremely common in web applications, it seems that it's integration in data mining software has lagged.
+After several unsuccessful attempts to view our data through WEKA, we opted for a different tool, RapidMiner.
+From here, we've taken two routes: 1) running a python script to convert the JSON to CSV and 2) import the data into a local MongoDB instance and connect via the proper port.
+Both methods have finally allowed some degree of visualization to better direct our inquiry.
+
+\section{Results}
+
+\subsection{Businesses}
+
+\quad After importing the CSV formatted version of the data we tested some basic visualization of the data.
+For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it?s located.
+Basic graphs like these help refine our question.
+For example, why does Illinois have significantly lower average stars than others?
+
+\quad The most pervasive issue with our business data is that many entries are missing data, most notably the 'attributes' and address information.
+Since attributes are managed by the account owner, they vary widely across the businesses.
+
+\begin{figure}
+\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
+\end{figure}
+
+\subsection{Users}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth, center]{u_reviews_vs_stars}
+\end{figure}
+
+\quad
 \section{Next Steps}

### 2017-04-11 code/csv_conversion/convert.py
Added source for convert.py

@@ -2,6 +2,7 @@
 Convert Yelp Academic Dataset from JSON to CSV
 Requires Pandas (https://pypi.python.org/pypi/pandas)
 By Paul Butler, No Rights Reserved
+source: https://gist.github.com/paulgb/5265767
 '''
 
 import json

### 2017-04-11 paper/samplebody-journals.tex
pdf update

--- 2017-04-11 paper/samplebody-journals.tex
@@
 
-\quad Our  rst objective is to determine if certain attributes are indica- tive of business success.
+\quad Our first objective is to determine if certain attributes are indicative of business success with respect to their average review rating (see sec.
+2.3 for revenue relation).
 If we can find a way to predict whether or not a business will do well later on in its operation, that information will be useful to businesses looking to improve their model.
 
-\quad Our second objective is to analyze the text of reviews for a sen- timent.
+\quad Our second objective is to analyze the text of reviews for a sentiment.
 The sentiment analysis will show the emotions conveyed by the reviewer as being pleased or angry at their experience with the business.
@@
 This information could also contribute to a business model.
-As an example, if certain loca- tions are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.
+As an example, if certain locations are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.
 
-\quad Our last objective is to  nd recurring problems in businesses.
-We will analyze review text to  nd the speci c issues that people are having with businesses.
+\quad Our last objective is to find recurring problems in businesses.
+We will analyze review text to find the specific issues that people are having with businesses.
 If a business has many reviews which all indicate a similar issue, then that is something a business owner would want to be aware of so that they can eliminate the problem.
@@
 From there they utilized various feature reduction and clustering techniques to build a prediction model.
-They were able to identify features that o ered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.
+They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.
 
@@
 (2013) utilized probabilistic models to discover under- lying subtopics within Yelp reviews.
-This o ered a framework to categorize the reviews based on review text.
+This offered a framework to categorize the reviews based on review text.
 From here, they were able to analyze the prevalence of various star ratings within each category.
-In this way, they could advise business owners as to the common trends among good and bad reviews that may not be ap- parent without considering the review text.
+In this way, they could advise business owners as to the common trends among good and bad reviews that may not be apparent without considering the review text.
 McAuley and Leskovec built on this idea by combining such latent review topics with hidden rating driving factors, allowing them construct a rating prediction model.
@@
 
-In one of the more in uential papers on the subject, Michael Luca showed a correlation between Yelp rating and revenue, highlighting the fact that rounding to half stars could signi cantly impact two similarly-rated restaurants if their averages fell on either side of the divide (e.g.
+In one of the more influential papers on the subject, Michael Luca showed a correlation between Yelp rating and revenue, highlighting the fact that rounding to half stars could significantly impact two similarly-rated restaurants if their averages fell on either side of the divide (e.g.
 3.24 and 3.26).
@@
 The dataset is inconsistent and cannot be relied upon for each business.
-For instance, one of these attributes is ?Bike Parking?.
+For instance, one of these attributes is 'Bike Parking'.
 This attribute is either true, false, or nonexistent.
-Speci cally the fact that this attribute might not be available for multiple businesses makes it unreliable.
+Speciifcally the fact that this attribute might not be available for multiple businesses makes it unreliable.
 
-\quad In addition, the data is not only restaurants but any kind of busi- ness.
-As a result they have another piece of data which is the cat- egories list.
+\quad In addition, the data is not only restaurants but any kind of business.
+As a result they have another piece of data which is the categories list.
 For each business this attribute attempts to classify the business.
-However some categories are extremely broad and the same type of business can be described in di erent ways.
+However some categories are extremely broad and the same type of business can be described in different ways.
 One example is that there is a bakery, but it?s also categorized as food.
-Many businesses will sell food, but the category of food will not be in their list of categories and instead it will be  lled with more speci c categories like Italian food.
+Many businesses will sell food, but the category of food will not be in their list of categories and instead it will be labelled with more specific categories like Italian food.
 
@@
 
-\quad The Yelp dataset contains up to one million di erent attributes about businesses from 11 di erent cities in 4 countries.
-The majority of these attributes will be unnecessary so there will be a large e ort in data reduction.
+\quad The Yelp dataset contains up to one million different attributes about businesses from 11 different cities in 4 countries.
+The majority of these attributes will be unnecessary so there will be a large effort in data reduction.
 For instance there are photographs included, but our questions does not require image analysis.
@@
 	\end{verbatim}
-	\item{Informationon144,000businessesin11citiesfrom4coun- tries (incl. the US, Germany, Canada, and the U.K. (with 1.1 million business attributes)}
+	\item{Information on 144,000 businesses in 11 cities from 4 countries (incl. the US, Germany, Canada, and the U.K. (with 1.1 million business attributes)}
 	\begin{verbatim}
@@
 Fortunately, there are a number of options, considering our educational status, that will allow large storage volume as well as low latency read/write access.
-Hosting in such a manner will requireA number of Python libraries like Dora and FTFY (?fixes text for you?) intended for data cleaning and preprocessing will be used for these purposes.
+Hosting in such a manner will require a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing will be used for these purposes.
 Yelp has already performed a substantial amount of cleaning on the data, so it is likely that minimal additional cleaning will be necessary, save for replacing missing values as mentioned above.
 
+\quad
+
 \quad Numpy and Scipy will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation.
-Finding angry reviews and recurring problems will make heavy use of NLP, so the team will rely on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, and possibly some custom supervised text classi ers trained with the Facebook fastText pretrained word vectors to perform sentiment analysis on the review text.
+Finding angry reviews and recurring problems will make heavy use of NLP, so the team will rely on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, and possibly some custom supervised text classifiers trained with the Facebook fastText pre-trained word vectors to perform sentiment analysis on the review text.
 
@@
 
+\section{Progress}
+
+\quad There have been several major changes to the project tools and methods.
+Initially we planned on using a relational database, but ran into a lot of trouble.
+The inconsistencies between the user and business data made it too difficult to successfully import it into the database, so instead we decided to import the data straight into Weka.
+However, the format of the JSON file did not allow a successful import into Weka.
+
+\quad Next we tried to convert the JSON file into CSV format. Once again Weka had issues due to the mismatch in attributes for each business (some had more attributes while others had less).
+
+\quad Instead we opted to use a different tool, RapidMiner.
+Like Weka, RapidMiner incorporates a extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining.
+RapidMiner is commercial software, but an educational version is available that removes limitations on the free version like only being able to process 10,000 rows at a time.
+Additionally, RapidMiner is more compatible with the missing data in our dataset and automatically corrects the formatting issues we had when attempting to use Weka.
+It does not support JSON so we reused the conversion code (found on the team GitHub as convert.py) to convert the data files to CSV.
+
+\section{Results}
+
+\subsection{Businesses}
+
+\quad After importing the CSV formatted version of the data we tested some basic visualization of the data.
+For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it?s located.
+Basic graphs like these help refine our question.
+For example, why does Illinois have significantly lower average stars than others?
+
+\quad The most pervasive issue with our business data is that many entries are missing data, most notably the 'attributes' and address information.
+Since these values are managed by the account owner, they vary widely across the businesses.
+
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
+\end{figure}
+
+\subsection{Users}
+
+\quad There are some very interesting features here. It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem.
+
+\begin{figure}[!h]
+\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
+\end{figure}
+
+\quad
 \section{Next Steps}
 
-\quad Some major milestones before Spring Break (March 27th) will include organizing and integrating all the data in a relational database, as well as cleaning and preprocessing the data.
-After these steps have been completed, it will be much easier to begin preliminary investigations into our main questions for the project, such as reducing the entire set of attributes into just those that will be mined for each question.
-Milestones after Spring Break prior to April 11th will include using the tools listed above to answer two of our objective questions, while work after that will focus on answering the remaining question and writing the final report.
+\quad It is still necessary to reduce dimensionality.
+We?ve currently been working on the whole dataset and simply ignoring all attributes that fall in the category ?other.?
+This is cumbersome and it will become increasingly detrimental as we use more advanced techniques like clustering.
+The most prioritized attribute is the attributes portion.
+This attribute has the most issues because it is a miscellaneous catch-all where any attribute that didn?t fit with the others is thrown in.
+The issue is that it totally lacks consistency in the number and types of miscellaneous attributes included for each business.
+One business may have a Free Wifi attribute while others may not.
+After looking through the data it was clear that there wasn?t one attribute in this misc.
+category that was reliably present for all businesses.
+Although we have moved on to work with the other attributes, this one still needs work to decide if it should be omitted completely or if default values should be set instead.
+
+\quad In addition, we need to combine the businesses and reviews dataset.
+This becomes extremely useful when examining the sentiment of a review and finding a correlation with the attributes of a business.
+This helps us build a model to potentially predict if a business receives negative or positive reviews based on its attributes.
+
+\quad As mentioned before, we will be interested in clustering businesses together as to whether they receive positive or negative reviews based on their attributes.
+This can be difficult with the high number of dimensions of these datasets.
+We will try CLIQUE and PROCLUS for subspace clustering.
+PROCLUS will be particularly useful because the subspace for one business to qualify as ?positively reviewed?
+could be different than the subspace for another business.
+This allows us to process this data faster, but also consider the complexities of various subspaces.
 

### 2017-04-11 code/nltk_review_subset_process.py
Adding file to find review sentiment with nltk

@@ -1,0 +1,41 @@
+from nltk.sentiment.vader import SentimentIntensityAnalyzer
+sid = SentimentIntensityAnalyzer()
+
+infile = "JSON/reviews/yelp_academic_dataset_review.json"
+# infile = "JSON/reviews/example.json"
+
+count = 0
+neg_count, pos_count = 0, 0
+avg_neg, avg_pos = 0, 0
+avg_all = 0
+
+with open(infile) as f:
+    for line in f:
+        if count >= 100000:
+            break
+        count += 1
+        ss = sid.polarity_scores(line[149:-48])
+
+        # for k in sorted(ss):
+        #     print k, ss[k],
+        # print
+
+        if ss['compound'] < -0.2:
+            neg_count += 1
+            avg_neg += ss['neg']
+        elif ss['compound'] > 0.2:
+            pos_count += 1
+            avg_pos += ss['pos']
+        if (count % 5000) == 0:
+            print "Processing", count, "of 4153150 (%5.3f %%)" % ((count/float(4153150)) * 100)
+        avg_all = avg_all + ss['pos'] - ss['neg']
+
+print
+print "Negative count: ", neg_count
+print "Average neg score among negative reviews: ", avg_neg / float(neg_count)
+print
+
+print "Positive count: ", pos_count
+print "Average pos score among positive reviews: ", avg_pos / float(pos_count)
+
+print "Total average: ", avg_all / float(count)

### 2017-04-11 paper/samplebody-journals.tex

@@
 
+Figure 2, depicts a graph of longitude vs. latitude vs. review count for businesses.
+
+\begin{figure}[!h]
+\includegraphics[width=0.75\textwidth, center]{3d_businesses_lat_long}
+\end{figure}
+
+In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect.
+However, it also shows visually the outliers.
+There are various businesses that have only a few reviews.
+In this context it is easier to identify outliers and also see that businesses concentrate in certain locations.
+Therefore we can reduce the dataset to only the cities that meet the threshold of businesses or total review counts.
+
 \subsection{Users}
 
-\quad There are some very interesting features here. It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem.
+\quad There are some very interesting features here.
+It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem.
+The figure below shows a plot of the user data.
+We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews.
+The most significant yet hardest to process attribute is the 'friends' list.
 
@@
 One business may have a Free Wifi attribute while others may not.
-After looking through the data it was clear that there wasn?t one attribute in this misc.
+After looking through the data it was clear that there wasn't one attribute in this misc.
 category that was reliably present for all businesses.

### 2017-04-11 paper/samplebody-journals.tex
added .gitignore, final pdf update

@@
 \quad Yelp is a website which allows users to create an account and review businesses.
-Our goal is to analyze data from Yelp to gain further insight for business owners to develop a satisfactory experience for their customers.
-Yelp has a public dataset available on their website with information about users, businesses, and reviews.
+Our goal is to analyze data from Yelp to gain further insight for business owners to develop a more satisfactory experience for their customers.
+Yelp has a large public dataset available on their website with information about users, businesses, and reviews.
 
@@
 2.3 for revenue relation).
-If we can find a way to predict whether or not a business will do well later on in its operation, that information will be useful to businesses looking to improve their model.
+For the businesses with many reviews, time-series data of the average star rating of the business mined from user reviews can be analyzed with ML models like Hidden Markov Models or recurrent neural networks in order to learn the likelihood of a business failing or succeeding in the future based on its current ratings.
+Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business's chances of success.
+Additionally, information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.
 
-\quad Our second objective is to analyze the text of reviews for a sentiment.
-The sentiment analysis will show the emotions conveyed by the reviewer as being pleased or angry at their experience with the business.
+\quad Another way sentiment analysis will prove helpful in this project is to analyze the emotions conveyed by the reviewer as being pleased or angry at their experience with the business.
 Using the results of this analysis we hope to be able to determine if certain times of day, city locations, or other factors contribute to the experience of a customer.
@@
 As an example, if certain locations are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.
+This type of information will be less useful to businesses, but it could prove very helpful for customers or advertisers to know that businesses in a certain area tend to be reviewed much higher or lower than the average business in cities of a comparable size.
 
 \quad Our last objective is to find recurring problems in businesses.
-We will analyze review text to find the specific issues that people are having with businesses.
-If a business has many reviews which all indicate a similar issue, then that is something a business owner would want to be aware of so that they can eliminate the problem.
+With natural language processing tools, we can analyze review text to find the specific issues that people are having with businesses via keywords or repeated references to specific aspects of the business.
+Again, this will be most helpful in analyzing reviews for businesses with many reviews.
+If one of these businesses has many reviews which all indicate a similar issue, then that is useful information that a business owner would want to be aware of so that they can eliminate the problem, if possible.
 
@@
 
-A previous winner, Hood et al., published a paper on assessing a business?s current rating/review state and future expectations with respect to review numbers on Yelp.
+A previous winner, Hood et al., published a paper on assessing a business's current rating/review state and future expectations with respect to review numbers on Yelp.
 The bulk of their work centered around the creation of additional features for the data (e.g.
@@
 
-The dataset has many optional attributes.
-There is an attribute called 'attributes' that is a list of optional attributes.
-The dataset is inconsistent and cannot be relied upon for each business.
-For instance, one of these attributes is 'Bike Parking'.
-This attribute is either true, false, or nonexistent.
-Speciifcally the fact that this attribute might not be available for multiple businesses makes it unreliable.
+\quad The dataset has many optional attributes.
+There is an attribute simply called 'attributes' that is a shopping list of optional fields for each business.
+This part of the dataset is inconsistent and the same fields cannot be relied upon to be present for each business.
+For instance, one of these attributes is 'Bike Parking.' This attribute is either true, false, or nonexistent.
+In the cases when it is nonexistent, analysis based upon this attribute is obviously unreliable.
+Most of these miscellaneous attributes are present for much less than a majority of the data, so any mining based upon them would have to first drastically trim the data available for analysis by eliminating more than half the data.
+For this reason, we have chosen to eliminate the miscellaneous category rather than most of our data.
 
-\quad In addition, the data is not only restaurants but any kind of business.
-As a result they have another piece of data which is the categories list.
-For each business this attribute attempts to classify the business.
-However some categories are extremely broad and the same type of business can be described in different ways.
-One example is that there is a bakery, but it?s also categorized as food.
-Many businesses will sell food, but the category of food will not be in their list of categories and instead it will be labelled with more specific categories like Italian food.
+\quad In addition, the data is not only restaurants but many kinds of businesses.
+A datum associated with each business attempts to classify the business as a member of a category, such as 'restaurants,' 'beaches,' or 'shopping.' However some categories are extremely broad and the same type of business can be described in different ways.
+One example is that a bakery may be classified as belonging to 'bakeries,' but also generically as 'food.' Conversely, many businesses sell food, but the generic category 'food' will not be in their list of categories and instead the business will be labelled with more specific categories like 'Italian food,' diminishing the usefulness of these categories.
 
-\quad The majority of the data cleaning process will be involved in handling missing data.
-It is possible to  ll in these missing values with default values like false but this can be dangerous to make assumptions about what businesses have and don?t have.
+\quad The dataset provided by the Yelp organizers is fairly clean already, and the majority of the data cleaning process will be involved in handling missing data.
+It is possible to fill in these missing values with default values like 'false', or 'none' but it can be dangerous to make assumptions about what businesses have and don't have - for example, filling in a 'gluten-free' attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.
 
-\quad The Yelp dataset contains up to one million different attributes about businesses from 11 different cities in 4 countries.
+\quad The Yelp dataset contains up to a million different attributes about businesses from 11 different cities in 4 countries.
 The majority of these attributes will be unnecessary so there will be a large effort in data reduction.
-For instance there are photographs included, but our questions does not require image analysis.
-There also attributes such as postal codes that are unnecessary because latitude and longitude is more accurate.
+For instance, the dataset includes photographs of the exterior and interior of some businesses (also many images of various kinds of food served by the businesses), but our questions does not require image analysis.
+There also redundant attributes such as postal codes that are unnecessary because latitude and longitude is more accurate.
 
@@
 
-The Yelp dataset comes from the Yelp Dataset Challenge.
+\quad The Yelp dataset comes from the Yelp Dataset Challenge.
 This is an ongoing challenge and this is the ninth round.
@@
 
-\quad The tools and methods used for this project will include a database for storing and integrating the data, libraries and modules for cleaning and preprocessing, and various tools for transformation, mining, pattern evaluation, and knowledge presentation.
-Given the json-formatted data, we've opted for using a NoSQL database like MongoDB or Cassandra.
-Fortunately, there are a number of options, considering our educational status, that will allow large storage volume as well as low latency read/write access.
-Hosting in such a manner will require a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing will be used for these purposes.
-Yelp has already performed a substantial amount of cleaning on the data, so it is likely that minimal additional cleaning will be necessary, save for replacing missing values as mentioned above.
+\quad The tools and methods used for this project will no longer include a database for storing or integrating the data.
+Due to difficulties described in section 6 and relating to the unknown JSON variant used by the preparers of the Yelp dataset, storing the data in a database is not feasible at this time.
+Given that the data is already quite clean save for the format, with minimal replacement of missing values and some other changes necessary, it is possible for each group member to have a copy of the data stored locally that may be processed with our tools, with smaller subsets of data, results, and scripts shared via GitHub.
 
-\quad
+\quad The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?' machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical.
+We will still use various Python libraries and modules for cleaning and preprocessing, and various other tools for transformation, mining, pattern evaluation, and knowledge presentation.
+The lack of a centralized database in removes the need for a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing, but increases the necessity of group communication to reduce duplication of work and miscommunications.
+We currently use and will continue to use Slack, email, and in-person meetings to communicate and synchronize our efforts.
 
-\quad Numpy and Scipy will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation.
-Finding angry reviews and recurring problems will make heavy use of NLP, so the team will rely on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, and possibly some custom supervised text classifiers trained with the Facebook fastText pre-trained word vectors to perform sentiment analysis on the review text.
+\quad Numpy, Scipy, and Pandas will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation.
+Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python's Natural Language Toolkit for sentiment analysis and other language processing.
 
@@
 
-\quad Evaluating the sentiment analysis of the review text will likely be simple given that we will be using tools like Weka and Microsoft Sentiment Analysis (see following section) to perform sentiment analysis rather than creating entirely custom algorithms.
+\quad Also, evaluating the sentiment analysis of the review text will likely be simple given that we will be using tools like NLTK (see following section) to perform sentiment analysis rather than creating entirely custom algorithms.
 These tools incorporate helpful metrics to analyze the error of a particular output.
@@
 For evaluating recurring problems, we will employ a similar approach, except we will group reviews by business and analyze them as a time series to and recurring issues.
+For recurring problems specifically, the most expedient method of evaluation may well be to read a small number of reviews and manually determine evaluate the performance of our tools in detecting recurrent issues.
 
-\quad For all three of the objective questions, employing association rule frameworks that use support-confidence frameworks will be helpful to and strong associations, while null-invariant measures like the Kulczynski measure or the cosine measure will be helpful to determine how interesting these associations are.
+\quad For all three of the objective questions, employing association rule frameworks that use support-confidence frameworks will be helpful to find and measure strong associations, while null-invariant measures like the Kulczynski measure or the cosine measure will be helpful to determine how interesting these associations are.
 
@@
 
-\quad There have been several major changes to the project tools and methods.
-Initially we planned on using a relational database, but ran into a lot of trouble.
-The inconsistencies between the user and business data made it too difficult to successfully import it into the database, so instead we decided to import the data straight into Weka.
-However, the format of the JSON file did not allow a successful import into Weka.
+\quad There have been several major changes to the project tools and methods since the project proposal.
+Initially we planned on using a relational database to store, integrate, and access the data, but but we encountered many difficulties in just uploading the data.
+The inconsistencies between the user and business data made it exceedingly difficult to successfully import the data en masse into the database, so instead we decided to import the data straight into Weka.
+However, the format of the JSON file also did not allow a successful import into Weka.
 
-\quad Next we tried to convert the JSON file into CSV format. Once again Weka had issues due to the mismatch in attributes for each business (some had more attributes while others had less).
+\quad Next we converted the JSON file into CSV format using a Python script and tried to upload it to Weka again.
+Once again Weka had issues due to the mismatch in attributes for each business (some had more attributes while others had less, likely due to the ?miscellaneous?
+category mentioned in section 3.1).
 
 \quad Instead we opted to use a different tool, RapidMiner.
-Like Weka, RapidMiner incorporates a extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining.
-RapidMiner is commercial software, but an educational version is available that removes limitations on the free version like only being able to process 10,000 rows at a time.
-Additionally, RapidMiner is more compatible with the missing data in our dataset and automatically corrects the formatting issues we had when attempting to use Weka.
+Like Weka, RapidMiner incorporates extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining.
+RapidMiner is commercial software, but an educational version is available that removes certain limitations on the free version like only being able to process 10,000 rows at a time.
+Additionally, RapidMiner is much more compatible with the missing data in our dataset and automatically corrects the formatting issues we had when attempting to use Weka.
 It does not support JSON so we reused the conversion code (found on the team GitHub as convert.py) to convert the data files to CSV.
@@
 \quad After importing the CSV formatted version of the data we tested some basic visualization of the data.
-For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it?s located.
+For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it's located.
 Basic graphs like these help refine our question.
@@
 \begin{figure}[h]
-\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
+\includegraphics[width=0.5\textwidth, center]{b_rating_by_state}
+\label{fig:Average star rating by state}
+\caption{Note the outliers IL and NTH}
 \end{figure}
@@
 \begin{figure}[!h]
-\includegraphics[width=0.75\textwidth, center]{3d_businesses_lat_long}
+\includegraphics[width=0.5\textwidth, center]{3d_businesses_lat_long}
+\caption{Clusters likely indicate areas of high population}
+\label{fig:Review count by latitude and longitude}
 \end{figure}
@@
 
+\subsection{Reviews}
+
+\quad One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time.
+However, given that we were unable to find a suitable workaround for our difficulties with Weka, we used sentiment analysis tools from the Natural Language Toolkit (NLTK) to analyze the review language instead.
+NLTK is a language-processing and analysis platform that uses Python, and a good replacement for the language processing capabilities of Weka.
+
+\quad Using a custom script (see GitHub; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews.
+The remaining 5737 reviews had composite sentiment scores within -0.2 < score < 0.2, so they were excluded from further analysis since they were neutral.
+This means that the average Yelp review (at least within the analyzed subset) is positive rather than negative or neutral in sentiment, perhaps indicating that Yelp users are more inclined to write reviews of businesses with which they had a good experience.
+Only around 2.5\% of the data is currently processed by the file, because processing the entire file is prohibitively expensive.
+
 \subsection{Users}
@@
 We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews.
-The most significant yet hardest to process attribute is the 'friends' list.
+In face, it appears that the user data skews so heavily in this grouping that all other measures
 
 \begin{figure}[!h]
-\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
+\includegraphics[width=0.5\textwidth, center]{u_reviews_vs_stars}
+  \caption{Note how the density increases between 3.25 and 4.25 starts}
+  \label{fig:Plot of review count vs average star rating}
 \end{figure}
 
+\quad One of the most significant yet hardest to process attributes is the 'friends' list.
+This attribute provides an opportunity to model our data as a graph.
+Such a network could provide very relevant clustering through sparseness measures and cutting in addition to accounting for user/business influence or finding deeper trends according to network connectivity.
 \quad
@@
 \quad It is still necessary to reduce dimensionality.
-We?ve currently been working on the whole dataset and simply ignoring all attributes that fall in the category ?other.?
-This is cumbersome and it will become increasingly detrimental as we use more advanced techniques like clustering.
+We've currently been working on the whole dataset and simply ignoring all attributes that fall in the category 'other.' This is cumbersome and it will become increasingly detrimental as we use more advanced techniques like clustering.
 The most prioritized attribute is the attributes portion.
@@
 One business may have a Free Wifi attribute while others may not.
-After looking through the data it was clear that there wasn't one attribute in this misc.
+After looking through the data it was clear that there wasn?t one attribute in this misc.
 category that was reliably present for all businesses.
@@
 This allows us to process this data faster, but also consider the complexities of various subspaces.
+
+\quad The user and review datasets are the largest we have.
+We would like to use these to find useful information about the people and the reviews they give.
+Because Yelp is a platform on which anybody can create an account and write a review, it is possible that not everybody is honest and fair in their reviews.
+Furthermore, different people may assign scores for different reasons.
+
+\quad Some useful attributes in the user data alone are the number of reviews that an individual has made, the length of time a user has been active on Yelp, and whether their reviews tend to be tagged as useful by other users.
+
+\quad One additional consideration is processing power.
+Our chosen tool seems to be very computationally intensive, so we will very likely need to use sampled data to test hypotheses then turn to python scripts to actually run over the entirety of our training, testing, and validations sets.
+Furthermore, our data is currently only linked through review, user, and business id pointers.
+While there?s a great deal to be learned from each collection individually, we will need to connect our disparate sets by these terms.
+
+\quad Currently, the review dataset is too large to process in detail.
+In the future, we will continue taking subset samples of a more reasonable size and processing those instead, as well as increasing the complexity of our analysis.
+Combining the review dataset with the business dataset (using the ?business\_id?
+field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews.
+We hope that additional advanced sentiment analysis will be useful in determining underlying factors that influence a business's success, as well as in identifying recurrent problems.
+
+\quad Once we have our sentiment analysis we can combine this data with our user data and business data to find patterns in the sentiment of reviews.
+We would like to compare the sentiment of reviews to the area they are coming from, the businesses they apply to, and whether reviews of a given sentiment are considered beneficial by other users.
+Using this information we can then narrow down the factors that are most influential to the score of a business in order to draw our final conclusions.
 

### 2017-04-11 code/csv_conversion/convert_users.py

@@ -1,29 +1,0 @@
-'''
-Convert Yelp Academic Dataset from JSON to CSV
-Requires Pandas (https://pypi.python.org/pypi/pandas)
-By Paul Butler, No Rights Reserved
-'''
-
-import json
-import pandas as pd
-from glob import glob
-
-def convert(x):
-    ''' Convert a json string to a flat python dictionary
-    which can be passed into Pandas. '''
-    ob = json.loads(x)
-    for k, v in ob.items():
-        if isinstance(v, list):
-            ob[k] = ','.join(v)
-        elif isinstance(v, dict):
-            for kk, vv in v.items():
-                ob['%s_%s' % (k, kk)] = vv
-            del ob[k]
-    return ob
-
-for json_filename in glob('*.json'):
-    csv_filename = '%s.csv' % json_filename[:-5]
-    print 'Converting %s to %s' % (json_filename, csv_filename)
-    df = pd.DataFrame([convert(line) for line in file(json_filename)])
-    df.to_csv(csv_filename, encoding='utf-8', index=False)
-

### 2017-04-30 code/problems_over_time.py
Code for tracking distribution of business problems over time

@@ -1,0 +1,160 @@
+"""
+download docs: CLA
+python
+import nltk
+nltk.download()
+select appropriate packages on GUI and download
+if no GUI "NLTK Downloader\n---------------\n\td) Download\tl) List...", then
+d vader_lexicon
+d stopwords
+q
+
+*********http://firstmonday.org/ojs/index.php/fm/article/view/4944/3863#p2
+"""
+
+# strip fillers, then can graph (scatter) VADER sentiment about business over time
+# stem, then analyze common collocations involving the words "issue", "problem", etc.
+
+import json
+from time import clock
+import os
+import nltk
+from string import punctuation
+import plotly
+import plotly.graph_objs as go
+from nltk import FreqDist
+from nltk.corpus import wordnet as wn
+import datetime
+
+
+# negative_sentiment = ["worst", "rude", "terrible", "horrible", "bad", "awful",
+#                       "disgusting", "bland", "tasteless", "gross", "mediocre",
+#                       "overpriced", "worse", "poor", "flavorless",
+#                       "disappoint"]  # added flavorless, disappoint
+# first_person_plural = ["we", "us", "our", "ourselves"]  # removed "ours"
+# third_person = ["she", "he", "her", "him"]
+# past_tense_verbs = ["was", "were", "asked", "told", "said", "did", "charged",
+#                     "waited", "left", "took"]
+# narrative_sequencers = ["after", "then"]
+# common_nouns = ["manager", "waitress", "waiter", "customer", "attitude",
+#                 "waste", "poisoning", "money", "bill", "minutes"]
+#                 # removed "customers"; redundant from stemming
+#
+# low_rating_words = [negative_sentiment, first_person_plural, third_person,
+#                     past_tense_verbs, narrative_sequencers, common_nouns]
+# low_rating = negative_sentiment + first_person_plural + third_person  \
+#              + past_tense_verbs + narrative_sequencers + common_nouns
+
+
+
+# categories: 1. service issue; 2. food issue; 3. price issue; 4. undetermined
+service_issue = ["rude",
+                 "asked", "told", "said", "did", "waited", "took",
+                 "after", "then",
+                 "manager", "waitress", "waiter", "attitude", "bill", "minutes"]
+food_issue = ["disgusting", "bland", "tasteless", "gross", "flavorless",
+              "poisoning"]
+price_issue = ["overpriced", "charged", "waste", "money", "bill"]
+undet_issue = ["worst", "terrible", "horrible", "bad", "awful", "mediocre",
[…]

### 2017-04-30 code/problems_over_time.py
Changed to plot # of reviews by month

@@ -4,16 +4,15 @@ python
 import nltk
 nltk.download()
 select appropriate packages on GUI and download
-if no GUI "NLTK Downloader\n---------------\n\td) Download\tl) List...", then
-d vader_lexicon
-d stopwords
-q
 
-*********http://firstmonday.org/ojs/index.php/fm/article/view/4944/3863#p2
+
+*******http://firstmonday.org/ojs/index.php/fm/article/view/4944/3863#p2
 """
 
 # strip fillers, then can graph (scatter) VADER sentiment about business over time
 # stem, then analyze common collocations involving the words "issue", "problem", etc.
+
+# need to normalize issue dist. by # of reviews received for particular period - week; month?
 
 import json
 from time import clock
@@ -63,8 +62,9 @@ undet_issue = ["worst", "terrible", "horrible", "bad", "awful", "mediocre",
                "customer"]
 issue_list = [service_issue, food_issue, price_issue, undet_issue]
 
-dates = []
+dates, rounded_dates = [], []
 service_list, food_list, price_list, undet_list = [], [], [], []
+month_buckets, counts = [], []
 
 
 def main():
@@ -79,11 +79,20 @@ def main():
     path = 'JSON/over_1000_reviews/'
     count = 0
     with open(path + str(os.listdir(path)[1])) as infile:
+
         for line in infile:
             service, food, price, undet = 0, 0, 0, 0
             obj = json.loads(line)
+            d = datetime.datetime.strptime(obj['date'], "%Y-%m-%d")
 
-            if obj['stars'] <= 3:
+            round_d = d.replace(day=1)
+            if round_d not in month_buckets:
+                month_buckets.append(round_d)
+                counts.append(1)
+            else:
+                counts[month_buckets.index(round_d)] += 1
+
+            if obj['stars'] <= 3:  # additional processing
                 count += 1
 
                 text = str_tuple(obj['text'].split())
@@ -107,7 +116,8 @@ def main():
 
                 total_issue_stems = service + food + price + undet
                 if total_issue_stems >= 5:  # threshold exceeded, calculate percentages
-                    dates.append(datetime.datetime.strptime(obj['date'], "%Y-%m-%d"))
+                    dates.append(d)
+
                     service_per = service / float(total_issue_stems)
[…]

### 2017-04-30 words_and_categories.py
Words and categories matching

@@ -1,0 +1,13 @@
+words = ["worst", "rude", "terrible", "horrible", "bad", "awful", "disgusting",
+         "bland", "tasteless", "gross", "mediocre", "overpriced", "worse",
+         "poor", "flavorless", "disappoint", "we", "us", "our", "ourselves",
+         "she", "he", "her", "him", "was", "were", "asked", "told", "said",
+         "did", "charged", "waited", "left", "took", "after", "then", "manager",
+         "waitress", "waiter", "customer", "attitude", "waste", "poisoning",
+         "money", "bill", "minutes"]
+
+# categories:
+service = []
+food = []
+price = []
+unresolved = []  # catchall for words that don't fit in other categories

### 2017-04-30 words_and_categories-dalton.py
Added my sorted words and categories.

@@ -1,0 +1,13 @@
+words = ["worst", "rude", "terrible", "horrible", "bad", "awful", "disgusting",
+         "bland", "tasteless", "gross", "mediocre", "overpriced", "worse",
+         "poor", "flavorless", "disappoint", "we", "us", "our", "ourselves",
+         "she", "he", "her", "him", "was", "were", "asked", "told", "said",
+         "did", "charged", "waited", "left", "took", "after", "then", "manager",
+         "waitress", "waiter", "customer", "attitude", "waste", "poisoning",
+         "money", "bill", "minutes"]
+
+# categories:
+service = ["rude", "she", "he", "her", "him", "asked", "told", "said", "did", "waited", "left", "took", "after", "manager", "waitress", "waiter", "customer", "attitude", "minutes"]
+food = ["disgusting", "bland", "tasteless", "gross", "flavorless", "poisoning"]
+price = ["overpriced", "charged", "money", "bill", "waste"]
+unresolved = ["worst", "terrible", "horrible", "bad", "awful", "mediocre", "worse", "poor", "disappoint", "we", "us", "our", "ourselves", "was", "were", "then"]  # catchall for words that don't fit in other categories

### 2017-05-01 code/save_reviews_to_business_files.py
Adding code to save reviews of frequently-reviewed businesses to separate files

@@ -1,0 +1,77 @@
+"""
+
+"""
+
+
+import json
+from time import clock
+import bisect
+
+businesses = 'JSON/businesses_many_reviews.json'
+reviews = 'JSON/yelp_academic_dataset_review.json'
+offset_b, offset_r = 0, 0
+business_list = []
+
+t1 = clock()
+b_count, r_count = 0, 0
+
+
+def index(a, x, yes, no):
+    # Locate the leftmost value exactly equal to x
+    i = bisect.bisect_left(a, x)
+    if i != len(a) and a[i] == x:
+        yes += 1
+    else:
+        no += 1
+    return yes, no
+
+with open(businesses) as b:
+
+    if offset_b:
+        b.seek(offset_b, 0)  # seek from beginning of the file to offset
+
+    for line in b:
+        obj = json.loads(line)
+        business = obj['business_id']
+        bisect.insort(business_list, str(business))
+        b_count += 1
+        # if b_count >= 5:
+        #     break
+    t2 = clock()
+    print "Done putting businesses in a list, time elapsed: " + str(t2 - t1)
+
+with open(reviews) as r:
+
+    if offset_r:
+        r.seek(offset_r, 0)
+
+    yes, no = 0, 0
+    for line in r:
+        try:
+            offset_r += len(line)
+            obj = json.loads(line)
+            business = obj['business_id']
+            old_yes = yes
+            yes, no = index(business_list, business, yes, no)
+            if old_yes + 1 == yes:  # write review to file keyed w/ business_id
+                business_file = 'JSON/over_1000_reviews/' + str(obj['business_id']) + '.json'
+                with open(business_file, "a") as f:
+                    f.write(line)
+
+            r_count += 1
+            if r_count % 100000 == 0:
+                print "processed " + str(r_count) + " of 4153150 ("  \
+                      + str(round((r_count/float(4153150))*100, 4)) + "%)"
+
+        except KeyboardInterrupt:
+            # can use offset_r to start processing at point in file where
+            # processing was terminated with f.seek(offset) in another run
+            print "\n" + str(offset_r)
+            break
+    print "yes, no: " + str(yes) + ", " + str(no)
+
+
+t3 = clock()
+print "Done matching business ids, total time elapsed: " + str(t3 - t1)
+
+print r_count, b_count

### 2017-05-01 paper/samplebody-journals.tex
added jupyter stuff:

@@
 
-In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect.
+\quad In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect.
 However, it also shows visually the outliers.
@@
 Therefore we can reduce the dataset to only the cities that meet the threshold of businesses or total review counts.
+
+\subsection{Business Categories}
+    \quad In an effort to garner more information about user behavior, we needed to augment the initial data with that from the businesses and reviews, the latter of which serves as the link between the former two.
+In doing so, we attempted to profile the most frequent business reviewed by each user.
+The two troublesome business features were the user-defined ?tags?
+and ?categories?.
+The categories proved to be a more manageable problem as they are selected from a finite set of options.
+With numbers of selected categories ranging from two to twenty-eight out of over 1000 possible choices, we sought to reduce this high degree of dimensionality.
+    \quad Sparse binary data of this kind quickly becomes unwieldy as traditional distance measures are highly prone to error.
+Such measures include K-Means, PCA analysis (for dimensionality reduction), DBSCAN, and Agglomerative Clustering.
+Typically, the computational cost is prohibitive (Plumpley, et.
+al., 2002) or the numerical values are highly unstable during computation.
+The goal in our application is to reduce the effective categorical space and classify the businesses into a narrower subset?on the order of ~50.
+Not surprisingly, this scenario is almost identical to that encountered when processing and analyzing large corpuses of text.
+Almost every form of textual representation falls into the ?bag-of-words?
+paradigm where vectors are used to represent each document, and each element corresponds to either the frequency of a given term in the dictionary (i.e.
+all words from all documents) or the binary existence of the word in the document.
+Since each document is highly unlikely to have an instance of all terms, the combination of these vectors results in a very large, very sparse matrix of n documents by p terms.
+    \quad Much research has been done in recent years on Latent Semantic Analysis (or Indexing).
+This utilizes the fact that a given matrix?s Singular Value Decomposition provides us with a large degree of useful data concerning the nature of the matrix, namely which vectors and rows carry more weight as well as those with less impact.
+Furthermore, it provides a model through which new data can be projected onto the reduced subspace.
+The table below shows a sample of the reduced vector space and the first ten terms in each of the new topics.
 

### 2017-05-01 paper/samplebody-journals.tex
images

@@
 For the businesses with many reviews, time-series data of the average star rating of the business mined from user reviews can be analyzed with ML models like Hidden Markov Models or recurrent neural networks in order to learn the likelihood of a business failing or succeeding in the future based on its current ratings.
-Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business's chances of success.
+Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business?s chances of success.
 Additionally, information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.
@@
 
-A previous winner, Hood et al., published a paper on assessing a business's current rating/review state and future expectations with respect to review numbers on Yelp.
+A previous winner, Hood et al., published a paper on assessing a business?s current rating/review state and future expectations with respect to review numbers on Yelp.
 The bulk of their work centered around the creation of additional features for the data (e.g.
@@
 They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.
+Our approach will be similar in the application of feature reduction to reduce the dimensions of that data and clustering to find cities with many businesses, as well as to group similar businesses together.
+However, we do not plan to create additional features like the number of nearby similar businesses.
 
@@
 Huang et al.
-(2013) utilized probabilistic models to discover under- lying subtopics within Yelp reviews.
+(2013) utilized probabilistic models to discover underlying subtopics within Yelp reviews.
 This offered a framework to categorize the reviews based on review text.
@@
 
-\quad The dataset has many optional attributes.
+The dataset has many optional attributes.
 There is an attribute simply called 'attributes' that is a shopping list of optional fields for each business.
@@
 \quad In addition, the data is not only restaurants but many kinds of businesses.
-A datum associated with each business attempts to classify the business as a member of a category, such as 'restaurants,' 'beaches,' or 'shopping.' However some categories are extremely broad and the same type of business can be described in different ways.
-One example is that a bakery may be classified as belonging to 'bakeries,' but also generically as 'food.' Conversely, many businesses sell food, but the generic category 'food' will not be in their list of categories and instead the business will be labelled with more specific categories like 'Italian food,' diminishing the usefulness of these categories.
+A datum associated with each business attempts to classify the business as a member of a category, such as ?restaurants,?
+?beaches,?
+or ?shopping.?
+However some categories are extremely broad and the same type of business can be described in different ways.
+One example is that a bakery may be classified as belonging to ?bakeries,?
+but also generically as ?food.?
+Conversely, many businesses sell food, but the generic category ?food?
+will not be in their list of categories and instead the business will be labelled with more specific categories like ?Italian food,?
+diminishing the usefulness of these categories.
 
 \quad The dataset provided by the Yelp organizers is fairly clean already, and the majority of the data cleaning process will be involved in handling missing data.
-It is possible to fill in these missing values with default values like 'false', or 'none' but it can be dangerous to make assumptions about what businesses have and don't have - for example, filling in a 'gluten-free' attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.
+It is possible to fill in these missing values with default values like ?false?, or ?none?
+but it can be dangerous to make assumptions about what businesses have and don?t have - for example, filling in a ?gluten-free?
+attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.
 
@@
 
-\quad The Yelp dataset comes from the Yelp Dataset Challenge.
-This is an ongoing challenge and this is the ninth round.
-The challenge offers \$5,000 to the winner and the  final submission is an academic paper.
+The Yelp dataset comes from the Yelp Dataset Challenge.
+This is an ongoing challenge and this is the ninth round, so the challenge is very well-established by this point.
+The challenge offers \$5,000 to the winner and the final submission takes the form of an academic paper.
 
@@
 	"compliment_type": (compliment count),
-		...(same for each compliment type)...
+		...(same for each complement type)...
 	"type":"user"
 	\end{verbatim}
-	\item{Over 200,000 user-submitted photographs, 947,000 tips, and aggregated check-in data}
+	\item{Over 200,000 user-submitted photographs, 947,000 tips, and aggregated check-in data for most businesses}
 \end{itemize}
@@
 Given that the data is already quite clean save for the format, with minimal replacement of missing values and some other changes necessary, it is possible for each group member to have a copy of the data stored locally that may be processed with our tools, with smaller subsets of data, results, and scripts shared via GitHub.
-
-\quad The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?' machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical.
+The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?s machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical.
 We will still use various Python libraries and modules for cleaning and preprocessing, and various other tools for transformation, mining, pattern evaluation, and knowledge presentation.
@@
 \quad Numpy, Scipy, and Pandas will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation.
-Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python's Natural Language Toolkit for sentiment analysis and other language processing.
+Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python?s Natural Language Toolkit for sentiment analysis and other language processing.
 
@@
 
-\quad Evaluating our predictions as to whether a business success will hinge on comparing the star rating (from one to  ve stars) and our predicted star rating based on the other attributes in the review, such as day of the week and month, season, or location.
+\quad Evaluating our predictions as to whether a business success will hinge on comparing the star rating (from one to five stars) and our predicted star rating based on the other attributes in the review, such as day of the week and month, season, or location.
 This could be done by training an ML algorithm like Kernel Ridge regression on 80\% of the data and testing it on the remaining 20\%, and then adjusting parameters to minimize the error between the actual star ratings and our predicted star ratings for the testing data.
@@
 
-\quad There have been several major changes to the project tools and methods since the project proposal.
+There have been several major changes to the project tools and methods since the project proposal.
 Initially we planned on using a relational database to store, integrate, and access the data, but but we encountered many difficulties in just uploading the data.
@@
 
-\quad After importing the CSV formatted version of the data we tested some basic visualization of the data.
-For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it's located.
-Basic graphs like these help refine our question.
-For example, why does Illinois have significantly lower average stars than others?
+After importing the CSV version of the data, we tested some basic visualization tools from RapidMiner.
+For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it is located.
+Basic graphs like these can help refine our questions.
+For example, why does Illinois have significantly lower average stars than other states or provinces, even those that border it?
 
-\quad The most pervasive issue with our business data is that many entries are missing data, most notably the 'attributes' and address information.
+\quad The most pervasive issue with our business data is that many entries are missing data, most notably the miscellaneous 'attributes' and address information.
 Since these values are managed by the account owner, they vary widely across the businesses.
+However, some of these data are redundant given the other fields, such as detailed latitude-longitude data for each business.
 
 \begin{figure}[h]
-\includegraphics[width=0.5\textwidth, center]{b_rating_by_state}
-\label{fig:Average star rating by state}
-\caption{Note the outliers IL and NTH}
+\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
 \end{figure}
 
-Figure 2, depicts a graph of longitude vs. latitude vs. review count for businesses.
+RapidMiner can also produce 3D scatterplots which is useful because we plan to use latitude and longitude to locate businesses, since some of the business address data is missing and cannot (and should not) be replaced with default values or other common values used to replace missing data.
+Figure 2 depicts a graph of longitude vs.
+latitude vs.
+review count for businesses.
 
 \begin{figure}[!h]
-\includegraphics[width=0.5\textwidth, center]{3d_businesses_lat_long}
-\caption{Clusters likely indicate areas of high population}
-\label{fig:Review count by latitude and longitude}
+\includegraphics[width=0.75\textwidth, center]{3d_businesses_lat_long}
 \end{figure}
 
-\quad In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect.
+In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect.
 However, it also shows visually the outliers.
 There are various businesses that have only a few reviews.
-In this context it is easier to identify outliers and also see that businesses concentrate in certain locations.
+In this context it is easier to identify outliers and also see that businesses concentrate in certain locations, as is expected given Yelp?s declaration that the dataset focuses on businesses in 11 cities.
 Therefore we can reduce the dataset to only the cities that meet the threshold of businesses or total review counts.
@@
 
-\quad One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time.
+One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time.
 However, given that we were unable to find a suitable workaround for our difficulties with Weka, we used sentiment analysis tools from the Natural Language Toolkit (NLTK) to analyze the review language instead.
@@
 
-\quad Using a custom script (see GitHub; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews.
+\quad Using a custom script (see GitHub\; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews.
 The remaining 5737 reviews had composite sentiment scores within -0.2 < score < 0.2, so they were excluded from further analysis since they were neutral.
@@
 
-\quad There are some very interesting features here.
-It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem.
+The users dataset essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem.
 The figure below shows a plot of the user data.
 We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews.
-In face, it appears that the user data skews so heavily in this grouping that all other measures
 
 \begin{figure}[!h]
-\includegraphics[width=0.5\textwidth, center]{u_reviews_vs_stars}
-  \caption{Note how the density increases between 3.25 and 4.25 starts}
-  \label{fig:Plot of review count vs average star rating}
+\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
 \end{figure}
@@
 Such a network could provide very relevant clustering through sparseness measures and cutting in addition to accounting for user/business influence or finding deeper trends according to network connectivity.
+However, given the magnitude of the dataset, modeling the ?friends?
+of each user as a complete graph will be computationally difficult.
+To combat this, we could limit our model to the first n people the graph builder encounters only, since connections between groups of people are unpredictable.
+
 \quad
@@
 In the future, we will continue taking subset samples of a more reasonable size and processing those instead, as well as increasing the complexity of our analysis.
-Combining the review dataset with the business dataset (using the ?business\_id?
-field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews.
+Combining the review dataset with the business dataset (using the 'business\_id' field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews.
 We hope that additional advanced sentiment analysis will be useful in determining underlying factors that influence a business's success, as well as in identifying recurrent problems.

### 2017-05-02 paper/paper.tex
New draft of final paper

\section{Abstract}

\quad The dataset from the ninth round of the Yelp Dataset Challenge was analyzed to determine which of the listed attributes are the most indicative of business success, as indicated by average star rating. Additionally, large quantities of review text were analyzed for sentiment and keywords originating from narratives relating to trauma to identify reasons for poor customer reviews. Whether or not the business was open or permanently shuttered at the time a review was written proved to be one of the best indicators of business category, more so than either latitude or longitude. Regardless of business location, the majority of the reviews were positive in sentiment, which correlates well with the average star rating of the reviews (around 3.7 also irrespective of location). No universal recurring problems were identified within the corpus, but time-series data of keywords from reviews from businesses with over 1000 reviews provided a method to visualize potential problems with businesses in one of several categories: service, food, price, or other.

\section{Introduction/Motivation}

\quad Yelp.com allows users to create an account and post public reviews of businesses. Our goal was to analyze the extensive data provided by Yelp as part of the Yelp Dataset Challenge [1] to gain further insight for business owners to develop a more satisfactory experience for their customers. Yelp has a large public dataset available on their website with information about users, businesses, and reviews.

\quad Our first objective was to determine if certain attributes are indicative of business success with respect to their average review rating (see Related Work, Reviews and Revenue). Decision trees and k-means clustering were employed to group businesses by attributes with the highest information gain or shortest distance in attribute space, respectively. Furthermore, sentiment analysis was used on the language present in the reviews and compared with the star reviews to provide more information on a business’s chances of success. Information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.

\quad Another way sentiment analysis proved helpful in this project was to analyze the emotions conveyed by the reviewer as being pleased or angry at their experience with the business. Using the results of this analysis, our goal was to determine how certain times of day, city locations, or other factors contribute to customer experience. This information could also contribute to a business model. As an example, if certain locations are more prone to angry customers, it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews. This type of information will be less useful to businesses, but it could prove very helpful for customers or advertisers to know that businesses in a certain area tend to be reviewed much higher or lower than the average business in cities of a comparable size. Also, to take an example, businesses reading their reviews on Yelp would be aware that the reviews in the area tend to be more negative than is average for the state, which in turn would prevent them from trying to fix problems that don’t exist.

\quad Our last objective was to find recurring problems in businesses. With natural language processing tools like NLTK, we analyzed review text to find the specific issues that people are having with businesses via keywords and repeated references to specific aspects of the business. If one of these businesses has many reviews which all indicate a similar issue, then that is useful information that a business owner would want to be aware of so that they can eliminate the problem, if possible. This was applied to reviews for businesses with a review count above a threshold, since businesses with few reviews are not likely to have enough information to examine recurring problems.

% Head 1
\section{Related Work}

\quad Since our project involves predictions surrounding businesses, we sought out similar research to guide our investigation. Thus, we focused our literature review on studies that utilized some form of sentiment analysis, constructed predictive models, and/or conducted meta-analysis of trends in the aforementioned sentiment, rating, or business interest.

% Head 2
\subsection{Predicting Business Attention}

\quad A previous winner, Hood et al. (2013) [2], published a paper on assessing a business’s current rating/review state and future expectations with respect to review numbers on Yelp. The bulk of their work centered around the creation of additional features for the data (e.g., the number of similar businesses within 1 km, features of subsets of reviews such as average rating and number of unique users, etc.). From there they utilized various feature reduction and clustering techniques to build a prediction model. They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as create a model for prediction that outperformed a basic linear regression. Our approach was similar in the application of feature reduction to reduce data dimensionality, and the use of clustering to not only find cities with many businesses, but also group similar businesses together. However, save for the total_compliments feature created to help analyze the user data, we did not add additional features to the dataset like the number of nearby similar businesses.

% Head 3
\subsection{Latent Subtopics}

\quad Huang et al. (2013) [3] utilized probabilistic models to discover underlying subtopics within Yelp reviews. This offered a framework to categorize the reviews based on review text. From here, they were able to analyze the prevalence of various star ratings within each category. In this way, they could advise business owners as to the common trends among good and bad reviews that may not be apparent without considering the review text. McAuley and Leskovec [4] built on this idea by combining such latent review topics with hidden rating driving factors, allowing them construct a rating prediction model.

\subsection{Reviews and Revenue}

\quad In one of the more influential papers on the subject, Michael Luca [5] showed a correlation between Yelp rating and revenue, highlighting the fact that rounding to half stars could significantly impact two similarly-rated restaurants if their averages fell on either side of the divide (e.g. 3.24 and 3.26). He goes on to build a model of market response based on review volume as well as the impact of user expertise. He concluded that such responses are consistent with a Bayesian learning model.

\subsection{Narratives and Consumer Sentiment}

\quad In an analysis of customer narratives embedded in Yelp reviews of restaurants in seven cities, Jurafsky et al. (2014) [6] determined that negative reviews are likely to incorporate words and other features previously associated with trauma. Specifically, the incidence of words and phrases with a negative emotional slant, words relating to actions performed by third-person actors like waiters or management, and first-person plural pronouns like “we” and “us” is higher for negative reviews than positive. Meanwhile, positive reviews were characterized by language associated with addiction and, in the case of more expensive restaurants, sensory pleasure.

\quad To determine precisely which words were statistically overrepresented in the category of negative reviews, in this case taken to be reviews with one star, the authors applied a “log-odds-ratio informative Dirichlet prior” (originally published in Monroe et al. (2008) [7]) to compare the frequency of each word in the full corpus to its frequency in the category of one-star reviews. The authors then sorted all the words by their log-odds association score to determine the 50 words that were most associated with one-star reviews. Rather than repeat their work, our group used a number of these words as given in the paper to determine the “issue distribution” of a particular restaurant over time (see Key Results, Tracking Problems).

\section{Dataset Information}

\quad The Yelp dataset comes from the Yelp Dataset Challenge. This is an ongoing challenge and this is the ninth round, so the challenge is very well-established by this point. The challenge offers \$5,000 to the winner and the final submission takes the form of an academic paper.

\quad We considered the following from the Yelp dataset:

\begin{itemize}
	\item{4.1 million reviews}
	\begin{verbatim}
	"review_id":(encrypted review id),
	"user_id":(encrypted user id),
	"business_id":(encrypted business id)"
	"stars":(star rating, rounded to half-stars),
	"date":(date, formatted: YYYY-MM-DD),
	"text":"review text",
	"useful":('useful' vote count),
	"funny":('funny' vote count),
	"cool": ('cool' vote count),
	"type": "review"
	\end{verbatim}
	\item{Information on 144,000 businesses in 11 cities from 4 countries (incl. the US, Germany, Canada, and the U.K. (with 1.1 million business attributes)}
	\begin{verbatim}
	"business_id":(encrypted business id),
	"name":"business name",
	"neighborhood": "hood name",
	"address":"full address",
	"city":"city",
	"state":"state"(if applicable),
	"postal code":"postal code",
	"latitude":latitude,
	"longitude":longitude,
	"stars":(star rating, rounded to half-stars),
	"review_count":(number of reviews),
	"is_open":0/1 (closed/open),
	"attributes":[(localized attribute tags)],
	"categories":[(localized category names)],
	"hours":[(hours strings)],
	"type": "business"
	\end{verbatim}
	\item{Over 1M users}
	\begin{verbatim}
	"user_id":(encrypted user id),
	"name":"first name",
	"review_count":(number of reviews),
	"yelping_since": (date, formatted: YYYY-MM-DD),
	"friends":[(encrypted user ids)],
	"useful":('useful' vote count sent by user),
	"funny":('funny' vote count sent by user),
	"cool": ('cool' vote count sent by user),
	"fans":"number of fans the user has",
	"elite":["an array of years the user was elite"],
	"average_stars":floating point average like 4.31,
	"compliment_type": (compliment count),
		...(same for each complement type)...
	"type":"user"
	\end{verbatim}
	\item{Over 200,000 user-submitted photographs, 947,000 tips, and aggregated check-in data for most businesses}
\end{itemize}

\section{Main Techniques}

\subsection{Data Cleaning and Preprocessing}

\quad Although the Yelp dataset was well formatted, its content was inconsistent. A lot of data was absent for nearly every attribute. Only a few attributes are guaranteed to exist for every business such as latitude and longitude. Businesses also contained an attributes attribute that had miscellaneous information about businesses; some examples of fields in attributes are Free Wifi, Garage Parking, etc. These are completely optional and are inconsistent among businesses. As a result of this inconsistency, we removed the attributes attribute from the dataset.

\quad The format of the data did not allow a successful import into Weka either in JSON or in CSV format, so instead we used RapidMiner, which incorporates extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining. RapidMiner is commercial software, but an educational version is available that removes certain limitations on the free version like only being able to process 10,000 rows at a time. Additionally, RapidMiner was much more compatible with the missing data in our dataset and automatically corrects some of the formatting issues we had when attempting to use Weka. It does not support JSON so we reused the conversion code (found on the team GitHub as convert.py) to convert the data files to CSV.

\quad The remaining missing data from other attributes such as categories or address were more rare than the attributes attribute. Instead of removing the attribute completely, they were autocompleted with default values if certain data values were non-existent. For example, the categories attribute would be set a default value of null if a business lacks a category. These cases are rare at between 200 and 300 per attribute which is small compared to the 144,000 businesses.

\quad In other cases only the subset of attributes were used to determine clusters using k-means. This because k-means fails to work on non-numerical values. So attributes such as address or name was completely avoided. In the case of k-means the attributes used were latitude, longitude, review count, and stars. These were the only numerical attributes that was provided in the businesses dataset.

\subsection{Decision Trees and K-means Clustering}

\quad The decision tree was used on the Yelp dataset to breakdown classification for the category of a business. The decision tree was used to help categorize businesses based on the existing categories attribute for each business. Using the categories section as a label helped create a decision tree which split mainly on the is_open, latitude, and longitude attribute. These attributes yielded the highest information gain. Since a decision tree requires a label, the categories section became most logical to classify a business. As a result the cleaning of the categories attribute was important because it required no missing values to successfully label all businesses.

\quad K-means clustering with k = 5 was also used to find clusters. In order to perform k-means it was necessary to have numeric attributes only. This would mean removing attributes like the address. In addition to this the dataset was also missing 67-166 values. Once again this is relatively low number so it was decided to fill in the missing values with the average of the appropriate columns.

\subsection{Correlation Calculations}

\quad In order to better understand the users of Yelp and draw more complete conclusions from our findings, we did a brief analysis on our user data. Each user has an average of the stars they give in their reviews and the number of different types of compliments they’ve received on their reviews. Users also have a number of fans. We first created a correlation matrix with RapidMiner in order to see the relationship between average stars and compliments. Following this a new feature was created called “total_compliments”, which was a summation of the compliments that a user has received. We then created a new correlation matrix to see what the overall relationship was between average stars and total compliments. Finally, we created another correlation matrix to examine the relationship between fans and total compliments. The correlation matrices were exported from RapidMiner to an image by first converting the matrix to an example data type using the ‘Converters’ extension and then writing that data to an excel file. The table can then be formatted in the excel file and exported to an image.

\subsection{Natural Language Processing}

\quad To investigate the linguistic characteristics of the review corpus, a good deal of preprocessing was necessary. At the beginning, all of the reviews were in one 3.46-GB file, which consistently crashed RapidMiner and necessitated the use of custom scripts focused on extracting and analyzing the information of interest. Since the review file was too large to open with RapidMiner, a different method of “joining” the business data and the review data was necessary. The script extract_businesses_many_reviews.py was used to locate businesses with more than 1000 reviews, and write the business names, IDs (a unique tag provided by Yelp), and other selected attributes like that business’s average star rating to a separate file. The number 1000 was chosen somewhat arbitrarily, but the following processing is the most informative for businesses with many reviews. There are 177 businesses with more than 1000 reviews in the dataset.

\quad The script save_reviews_to_business_files.py was used to save all reviews for all businesses in the commonly-reviewed list to separate files, such that each file, named <business_id>.json, contained all the reviews for the business with that business ID. This step was important in building the processing pipeline, but could easily be eliminated now that the pipeline has been developed, so reviews for a certain business could be identified while the script loops through the large review file and then processed without an extra “save” step in between.

\quad Once all reviews for a business were in one file, the script problems_over_time.py was used to analyze the reviews and graph the results. To accomplish this, each review in a <business_id>.json file with a star rating of 3 or fewer stars was tokenized, stemmed, and compared to the following list of words associated with traumatic narratives in Fig.4 on the left:

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{table_trauma_keywords}
\end{figure}

\quad For comparison, the list of 50 words identified by Jurafsky et al. (2014) is on the right (see Related Work, Narratives and Consumer Sentiment). Notably, none of the seven cities used in the paper were among the set of cities provided by the Yelp Dataset Challenge, so there was a guaranteed overlap of zero reviews between the two corpora. Also, the list of words used for problem tracking over time was stemmed, eliminating the need to include plurals and different forms like “managers” as well as “manager,” or “waiting” and “wait” as well as “waited.” The count of stemmed forms of each of the words in the above left table was normalized by the total number of reviews for that month and graphed over time.

\section{Key Results}

\subsection{Basic Statistics and RapidMiner}

\quad After importing the CSV version of the data into RapidMiner, we tested some basic visualization tools included with the package. For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it is located. Basic graphs like these can help refine our questions. For example, why does Illinois have significantly lower average stars than other states or provinces, even those that border it?

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
\end{figure}

\quad RapidMiner can also produce 3D scatterplots which is useful because we used latitude and longitude in conjunction with the city and state to locate businesses, since some of the business address data is missing and cannot be filled with default values or other common values used to replace missing data. Figure 2 depicts a graph of longitude vs. latitude vs. review count for businesses.

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth, center]{3d_businesses_lat_long}
\end{figure}

\quad In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expected. However, it also shows visually the outliers. There are various businesses that have only a few reviews. In this context it is easier to identify outliers and also see that businesses concentrate in certain locations, as is expected given Yelp?s declaration that the dataset focuses on businesses in 11 cities. Therefore we reduced the dataset to only the cities that meet the threshold of businesses or total review counts.

\subsection{Business Categories}

\quad In an effort to garner more information about user behavior, we needed to augment the initial data with that from the businesses and reviews, the latter of which serves as the link between the former two. In doing so, we attempted to profile the most frequent business reviewed by each user. The two troublesome business features were the user-defined ?tags? and ?categories?. The categories proved to be a more manageable problem as they are selected from a finite set of options. With numbers of selected categories ranging from two to 28 out of over 1000 possible choices, we sought to reduce this high degree of dimensionality.

\quad Sparse binary data of this kind quickly becomes unwieldy as traditional distance measures are highly prone to error. Such measures include k-means, PCA analysis (for dimensionality reduction), DBSCAN, and agglomerative clustering. Typically, the computational cost is prohibitive (Plumpley, et. al. (2002)) or the numerical values are highly unstable during computation. The goal in our application is to reduce the effective categorical space and classify the businesses into a narrower subset?on the order of ~50. Not surprisingly, this scenario is almost identical to that encountered when processing and analyzing large corpuses of text. Almost every form of textual representation falls into the ?bag-of-words? paradigm where vectors are used to represent each document, and each element corresponds to either the frequency of a given term in the dictionary (i.e., all words from all documents) or the binary existence of the word in the document. Since each document is highly unlikely to have an instance of all terms, the combination of these vectors results in a very large, very sparse matrix of n documents by p terms.

\quad Much research has been done in recent years on Latent Semantic Analysis (or Indexing). This utilizes the fact that a given matrix?s Singular Value Decomposition provides us with a large degree of useful data concerning the nature of the matrix, namely which vectors and rows carry more weight as well as those with less impact. Furthermore, it provides a model through which new data can be projected onto the reduced subspace. The table below shows a sample of the reduced vector space and the first ten terms in each of the new topics.

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth, center]{topics}
\end{figure}

\subsection{Decision Trees}

\quad The decision tree shows the separation of businesses by categories. The tree provided insight into where the categories are dominant. As seen in Fig. 5 the tree breaks down into ranges of different latitude and longitude. The location of business has a strong correlation with the type of business. For instance in Fig. 5 the tree shows that a latitude of greater than 43.784 and a longitude of less than or equal to -73.477 yielded a large amount of categories. This is because this coordinate is in between New York City and Montreal which are high dense cities. The decision tree shows the best paths to take to determine the category of a business. The tree relies entirely on three attributes to determine a business’s category, is_open, latitude, and longitude. The location of a business is important so it is logical that a more dense variety of categories appear between Montreal and New York City as opposed to Arizona where there are fewer categories.

\quad The more interesting finding was that is_open is one of the main contributors to the category of a business. In retrospect the connection is clear because the failure of a business can have a domino effect. It is common to see a plaza or circle of businesses close down together because the location lacks popularity. Interestingly the tree prioritizes the is_open attribute higher than either the latitude or the longitude. If a business is open or closed permanently can actually determine the location and category of the business. One way of interpreting this is that it is very likely for a business to be a certain category if it is open or closed. The location is reflective of the trends and needs of that community. As a result many businesses of that type will either thrive or close in a certain latitude and longitude. Using the Montreal example before, the Canadian city has influence from New York City below it, the French, and of course Canadian culture itself. This blend of culture is well reflected in the culinary culture of Montreal. The tree in Fig. 5 portrays this area having categories of Food Delis, Comfort Food, Restaurants, Creperies, and Desserts. If a business is open and falls within those coordinates then it is likely that the business serves food of some sort.

\quad By contrast, the closed businesses tend to be more specialized and have an exact purpose within a certain community. Their locations are in less populated cities. For instance in Fig. 6, businesses that are categorized as Heating and Air Conditioning is in a less populated area. The color key below the node indicates the ratio of competing businesses in that same area. In the case of Heating and Air Conditioning it is the monopoly in its area. The business specializes in Heating and Air Conditioning where only a single type of this business is necessary in a smaller community. It is also follows that these locations are more obscure because they exist in areas where changes in weather are more extreme requiring the aid of this business.

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{decision_tree_open}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{decision_tree_closed}
\end{figure}

\quad The use of k-means returned results that were surprisingly consistent in star reviews (Fig. 7). Many of the businesses are in highly populated cities. As a result the k-means struggled to separate these businesses by location. However, k-means reveals the significant difference in average review counts by location. For instance the average review count for cluster 0 is ~207 reviews compared to cluster 1 which has ~18 reviews on average. However the average star rating is roughly similar around 3.7 stars. This pattern is seen in the other three clusters where differences in review counts still result in similar average star ratings. Across all clusters the stars are all near 3.7 stars average reviews. In cluster 2 the average star rating is ~3.5 which varies more significantly from the other clusters. The responding latitude and longitude is in the Massachusetts and New Hampshire area. The east coast has significantly less average review, but also lower star ratings as seen before. This may indicate more strict culinary culture on the east coast compared to the west with a relaxed comfort food culture. Yelp is a Silicon Valley company so its adoption and usage is likely more significant in western states.

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{k_means_centroids}
\end{figure}

\subsection{User Behavior}

\quad The users dataset essentially provides a window into users' status, connections, and interactions within the Yelp ecosystem. The figure below shows a plot of the user density versus the average star rating. We see that the density grows around 3.25-4.25 stars with some traces of a bell curve regarding the number of reviews the user has written.

\begin{figure}[!h]
\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
\end{figure}

\quad Our correlation matrices for the user data showed some insight into the social behavior on Yelp. Our first matrix included each complement type and the average stars a user gave in their reviews (Fig. 8). This matrix showed that there was not a correlation between the number of stars a user would give in their reviews and the compliments they receive. This seems to indicate that users would not be motivated to give a business a better or worse review in order to gain favorability in their profile.

\quad The second matrix (Fig. 9) includes the average stars feature and the total compliments feature. Again, there does not seem to be a relationship between star ratings given and the compliments received.

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{correlation_matrix}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{star_and_compliment_matrix}
\end{figure}

\subsection{Tracking Problems}

\quad To answer the question of what types of problems were repeatedly present in a business, we built a utility that analyzes the language found in reviews. As mentioned in the Related Work section, Jurafsky et al. (2014) used a log-odds-ratio informative Dirichlet prior to determine the 50 words most statistically overrepresented in Yelp reviews with one star. We then used the stemmed forms of most of these words (see Fig. 4 in Main Techniques, Natural Language Processing) to determine the “issue distribution” for a particular business during a particular month, two examples of which are below as Fig. 10:

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{nltk_tracking}
\end{figure}

\quad The high variability in weight seen in from 2008-2010 for Earl of Sandwich is due to a higher prevalence of negative words during every other month, which was a common occurrence in many of the analyzed restaurants. Restaurants were the only type of business from the review data that this analysis was applied to, because the “training set” of words provided by Jurafsky et al. (2014) were obtained from an analysis of restaurants. The spikes every other month could be the result of customers with high hopes visiting and coming away disappointed, leaving waves of more positive and more negative reviews. Additionally, it is worth noting that the average star rating for Bachi Burger within the dataset is 4.0, and the corresponding value for Earl of Sandwich is 4.5. Although both businesses had greater than 2700 reviews in the dataset, the number with 3 stars or fewer was substantially fewer, limiting the effectiveness of the analysis and providing another mechanism for the spikes seen in both graphs.

\section{Applications}

\quad The decision tree provides insight into the separation of businesses by their category. This can help classify the category of a new business or new areas. For instance, given a new business located at x latitude and y longitude, Yelp can make predictions based on surrounding categories. The tree also reveals the significance of a business being open or permanently closed, and potential declines in certain communities of businesses. Communities that have closing businesses are likely to have surrounding businesses close as well. In terms of businesses, this can help entrepreneurs choose areas that aren’t in decline when starting a new business.

\quad Latitude and longitude are also significant to the category of a business. For businesses this is important because certain categories become abundant in certain areas. One example is that in both New York City and Montreal there are a significant number of creperies because of New York City’s large population density and Montreal’s French influences. New businesses should consider other locations where creperies aren’t as popular. This can provide opportunities for new or existing businesses to venture into new areas. It can also help reveal the niche in certain communities such as in Fig. 4 where certain areas have relatively high concentrations of fitness-related businesses and even martial arts. New businesses can add to concentrations of businesses with an extra twist or bring quality to a category. This quality can be better determined with the k-means clusters.

\quad The clusters determined by k-means portray where Yelp is more impactful to a business. West coast areas receive more reviews on average. This can be helpful for businesses to measure their performance. Businesses that want to receive more feedback may consider the clusters that have higher review counts. It also reveals a more active community of Yelp which could be better business. There may be more competition in this clusters, but success can result in exponential growth in an active Yelp community.

\quad As mentioned in the introduction, the natural language processing performed on the review text to identify negative sentiment and recurring problems can be leveraged by managers and other staff members to improve their ratings, thereby improving revenue.

\section{References}

\begin{enumerate}
	\item Yelp Dataset Challenge, Round 9: https://www.yelp.com/dataset_challenge

	\item Hood, B., Hwang, V., and King, J. Inferring future business attention. in The Yelp Challenge, (Carnegie Mellon University, Pittsburgh, PA, 2013), Yelp.com. Retrieved March 3, 2017, from Yelp.com: https://www.yelp.com/html/pdf/YelpDatasetChallengeWinner_InferringFuture.pdf.

    \item Huang, J., Rogers, S., and Joo, E. Improving restaurants by extracting subtopics from Yelp reviews. in iConference, (Berlin, Germany, 2014), Yelp.com. Retrieved March 3, 2017, from Yelp.com: https://www.yelp.com/html/pdf/YelpDatasetChallengeWinner_ImprovingRestaurants.pdf.

    \item McAuley, J., and Leskovec, J. Hidden factors and hidden topics: Understanding rating dimensions with review text. in 7th ACM Conference on Recommender Systems (Hong Kong, 2013), the ACM Conference Series on Recommenders Series. Retrieved March 3, 2017, from the Stanford Computer Science Department: https://cs.stanford.edu/people/jure/pubs/reviews-recsys13.pdf.

    \item Luca, M. Reviews, reputation, and revenue: The case of Yelp.com. Submitted to American Economic Journal: Applied Economics. Retrieved March 3, 2017, from Harvard Business School: http://www.hbs.edu/faculty/Publication%20Files/12-016_a7e4a5a2-03f9-490d-b093-8f951238dba2.pdf.

    \item Jurafsky, D., Chahuneau, V., Routledge, B.R., and Smith, N.A. Narrative framing of consumer sentiment in online restaurant reviews. First Monday, 19 (4). Retrieved April 29, 2017, from First Monday: http://firstmonday.org/ojs/index.php/fm/article/view/4944/3863.

    \item Monroe, B., Colaresi, M., and Quinn, K. Fightin’ words: Lexical feature selection and evaluation for identifying the content of political conflict. Political Analysis, 16 (4). 372-403.

\end{enumerate}

% Start of "Sample References" section

%\section{Typical references in new ACM Reference Format}
%A paginated journal article \cite{Abril07}, an enumerated
%journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
%a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
%\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
%followed by the same example, however we only output the series if the volume number is given
%\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
%a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
%in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
%an article in a proceedings (of a conference, symposium, workshop for example)
%(paginated proceedings article) \cite{Andler79}, a proceedings article
%with all possible elements \cite{Smith10}, an example of an enumerated
%proceedings article \cite{VanGundy07},
%an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
%a master's thesis: \cite{anisi03}, an online document / world wide web
%resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
%and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
%work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
%\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
%'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
%Boris / Barbara Beeton: multi-volume works as books
%\cite{MR781536} and \cite{MR781537}.
%
%A couple of citations with DOIs: \cite{2004:ITE:1009386.1010128,
%  Kirschmer:2010:AEI:1958016.1958018}.

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

### 2017-05-02 paper/paper.tex
Fixed figure labeling order

@@
 \quad Once all reviews for a business were in one file, the script problems_over_time.py was used to analyze the reviews and graph the results.
-To accomplish this, each review in a <business_id>.json file with a star rating of 3 or fewer stars was tokenized, stemmed, and compared to the following list of words associated with traumatic narratives in Fig.4 on the left:
+To accomplish this, each review in a <business_id>.json file with a star rating of 3 or fewer stars was tokenized, stemmed, and compared to the following list of words associated with traumatic narratives in Fig.
+1 on the left:
 
@@
 \quad After importing the CSV version of the data into RapidMiner, we tested some basic visualization tools included with the package.
-For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it is located.
+For instance, in Fig.
+2 we created a histogram of the average star for a business versus the state in which it is located.
 Basic graphs like these can help refine our questions.
@@
 \quad RapidMiner can also produce 3D scatterplots which is useful because we used latitude and longitude in conjunction with the city and state to locate businesses, since some of the business address data is missing and cannot be filled with default values or other common values used to replace missing data.
-Figure 2 depicts a graph of longitude vs.
+Fig.
+3 depicts a graph of longitude vs.
 latitude vs.
@@
 Furthermore, it provides a model through which new data can be projected onto the reduced subspace.
-The table below shows a sample of the reduced vector space and the first ten terms in each of the new topics.
+The table below as Fig.
+4 shows a sample of the reduced vector space and the first ten terms in each of the new topics.
 
@@
 
-\quad The users dataset essentially provides a window into users' status, connections, and interactions within the Yelp ecosystem.
-The figure below shows a plot of the user density versus the average star rating.
-We see that the density grows around 3.25-4.25 stars with some traces of a bell curve regarding the number of reviews the user has written.
-
-\begin{figure}[!h]
-\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
-\end{figure}
-
 \quad Our correlation matrices for the user data showed some insight into the social behavior on Yelp.
@@
 We then used the stemmed forms of most of these words (see Fig.
-4 in Main Techniques, Natural Language Processing) to determine the “issue distribution” for a particular business during a particular month, two examples of which are below as Fig.
+1 in Main Techniques, Natural Language Processing) to determine the “issue distribution” for a particular business during a particular month, two examples of which are below as Fig.
 10:
@@
 It can also help reveal the niche in certain communities such as in Fig.
-4 where certain areas have relatively high concentrations of fitness-related businesses and even martial arts.
+5 where certain areas have relatively high concentrations of fitness-related businesses and even martial arts.
 New businesses can add to concentrations of businesses with an extra twist or bring quality to a category.

### 2017-05-02 paper/samplebody-journals.tex
stuff

@@
+\section{Abstract}
+
+\quad The dataset from the ninth round of the Yelp Dataset Challenge was analyzed to determine which of the listed attributes are the most indicative of business success, as indicated by average star rating.
+Additionally, large quantities of review text were analyzed for sentiment and keywords originating from narratives relating to trauma to identify reasons for poor customer reviews.
+Whether or not the business was open or permanently shuttered at the time a review was written proved to be one of the best indicators of business category, more so than either latitude or longitude.
+Regardless of business location, the majority of the reviews were positive in sentiment, which correlates well with the average star rating of the reviews (around 3.7 also irrespective of location).
+No universal recurring problems were identified within the corpus, but time-series data of keywords from reviews from businesses with over 1000 reviews provided a method to visualize potential problems with businesses in one of several categories: service, food, price, or other.
+
 \section{Introduction/Motivation}
 
-\quad Yelp is a website which allows users to create an account and review businesses.
-Our goal is to analyze data from Yelp to gain further insight for business owners to develop a more satisfactory experience for their customers.
+\quad Yelp.com allows users to create an account and post public reviews of businesses.
+Our goal was to analyze the extensive data provided by Yelp as part of the Yelp Dataset Challenge to gain further insight for business owners to develop a more satisfactory experience for their customers.
 Yelp has a large public dataset available on their website with information about users, businesses, and reviews.
 
-\quad Our first objective is to determine if certain attributes are indicative of business success with respect to their average review rating (see sec.
-2.3 for revenue relation).
-For the businesses with many reviews, time-series data of the average star rating of the business mined from user reviews can be analyzed with ML models like Hidden Markov Models or recurrent neural networks in order to learn the likelihood of a business failing or succeeding in the future based on its current ratings.
-Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business?s chances of success.
-Additionally, information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.
+\quad Our first objective was to determine if certain attributes are indicative of business success with respect to their average review rating (see Related Work, Reviews and Revenue).
+Decision trees and k-means clustering were employed to group businesses by attributes with the highest information gain or shortest distance in attribute space, respectively.
+Furthermore, sentiment analysis was used on the language present in the reviews and compared with the star reviews to provide more information on a business’s chances of success.
+Information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.
 
-\quad Another way sentiment analysis will prove helpful in this project is to analyze the emotions conveyed by the reviewer as being pleased or angry at their experience with the business.
-Using the results of this analysis we hope to be able to determine if certain times of day, city locations, or other factors contribute to the experience of a customer.
+\quad Another way sentiment analysis proved helpful in this project was to analyze the emotions conveyed by the reviewer as being pleased or angry at their experience with the business.
+Using the results of this analysis, our goal was to determine how certain times of day, city locations, or other factors contribute to customer experience.
 This information could also contribute to a business model.
-As an example, if certain locations are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.
+As an example, if certain locations are more prone to angry customers, it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.
 This type of information will be less useful to businesses, but it could prove very helpful for customers or advertisers to know that businesses in a certain area tend to be reviewed much higher or lower than the average business in cities of a comparable size.
+Also, to take an example, businesses reading their reviews on Yelp would be aware that the reviews in the area tend to be more negative than is average for the state, which in turn would prevent them from trying to fix problems that don’t exist.
 
-\quad Our last objective is to find recurring problems in businesses.
-With natural language processing tools, we can analyze review text to find the specific issues that people are having with businesses via keywords or repeated references to specific aspects of the business.
-Again, this will be most helpful in analyzing reviews for businesses with many reviews.
+\quad Our last objective was to find recurring problems in businesses.
+With natural language processing tools like NLTK, we analyzed review text to find the specific issues that people are having with businesses via keywords and repeated references to specific aspects of the business.
 If one of these businesses has many reviews which all indicate a similar issue, then that is useful information that a business owner would want to be aware of so that they can eliminate the problem, if possible.
+This was applied to reviews for businesses with a review count above a threshold, since businesses with few reviews are not likely to have enough information to examine recurring problems.
 
 % Head 1
-\section{Prior Work}
+\section{Related Work}
 
-Since our problem involves predictions surrounding businesses, we sought out similar research to guide our investigation.
-Thus, we focus our literature review on studies that utilized some form of sentiment analysis, constructed predictive models, and/or conducted meta-analysis of trends in the aforementioned sentiment, rating, or business interest.
+\quad Since our project involves predictions surrounding businesses, we sought out similar research to guide our investigation.
+Thus, we focused our literature review on studies that utilized some form of sentiment analysis, constructed predictive models, and/or conducted meta-analysis of trends in the aforementioned sentiment, rating, or business interest.
 
@@
 
-A previous winner, Hood et al., published a paper on assessing a business?s current rating/review state and future expectations with respect to review numbers on Yelp.
-The bulk of their work centered around the creation of additional features for the data (e.g.
-number of similar businesses within 1km, features of subsets of reviews such as average rating and number of unique users).
+\quad A previous winner, Hood et al.
+(2013), published a paper on assessing a business’s current rating/review state and future expectations with respect to review numbers on Yelp.
+The bulk of their work centered around the creation of additional features for the data (e.g., the number of similar businesses within 1 km, features of subsets of reviews such as average rating and number of unique users, etc.).
 From there they utilized various feature reduction and clustering techniques to build a prediction model.
-They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.
-Our approach will be similar in the application of feature reduction to reduce the dimensions of that data and clustering to find cities with many businesses, as well as to group similar businesses together.
-However, we do not plan to create additional features like the number of nearby similar businesses.
+They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as create a model for prediction that outperformed a basic linear regression.
+Our approach was similar in the application of feature reduction to reduce data dimensionality, and the use of clustering to not only find cities with many businesses, but also group similar businesses together.
+However, save for the total_compliments feature created to help analyze the user data, we did not add additional features to the dataset like the number of nearby similar businesses.
 
@@
 
-Huang et al.
+\quad Huang et al.
 (2013) utilized probabilistic models to discover underlying subtopics within Yelp reviews.
@@
 
-In one of the more influential papers on the subject, Michael Luca showed a correlation between Yelp rating and revenue, highlighting the fact that rounding to half stars could significantly impact two similarly-rated restaurants if their averages fell on either side of the divide (e.g.
+\quad In one of the more influential papers on the subject, Michael Luca showed a correlation between Yelp rating and revenue, highlighting the fact that rounding to half stars could significantly impact two similarly-rated restaurants if their averages fell on either side of the divide (e.g.
 3.24 and 3.26).
@@
 
-\section{Proposed Work and Dataset Information}
+\subsection{Narratives and Consumer Sentiment}
 
-\subsection{Proposed Work}
+\quad In an analysis of customer narratives embedded in Yelp reviews of restaurants in seven cities, Jurafsky et al.
+(2014) determined that negative reviews are likely to incorporate words and other features previously associated with trauma.
+Specifically, the incidence of words and phrases with a negative emotional slant, words relating to actions performed by third-person actors like waiters or management, and first-person plural pronouns like “we” and “us” is higher for negative reviews than positive.
+Meanwhile, positive reviews were characterized by language associated with addiction and, in the case of more expensive restaurants, sensory pleasure.
 
-The dataset has many optional attributes.
-There is an attribute simply called 'attributes' that is a shopping list of optional fields for each business.
-This part of the dataset is inconsistent and the same fields cannot be relied upon to be present for each business.
-For instance, one of these attributes is 'Bike Parking.' This attribute is either true, false, or nonexistent.
-In the cases when it is nonexistent, analysis based upon this attribute is obviously unreliable.
-Most of these miscellaneous attributes are present for much less than a majority of the data, so any mining based upon them would have to first drastically trim the data available for analysis by eliminating more than half the data.
-For this reason, we have chosen to eliminate the miscellaneous category rather than most of our data.
+\quad To determine precisely which words were statistically overrepresented in the category of negative reviews, in this case taken to be reviews with one star, the authors applied a “log-odds-ratio informative Dirichlet prior” (originally published in Monroe et al.
+(2008)) to compare the frequency of each word in the full corpus to its frequency in the category of one-star reviews.
+The authors then sorted all the words by their log-odds association score to determine the 50 words that were most associated with one-star reviews.
+Rather than repeat their work, our group used a number of these words as given in the paper to determine the “issue distribution” of a particular restaurant over time (see Key Results, Tracking Problems).
 
-\quad In addition, the data is not only restaurants but many kinds of businesses.
-A datum associated with each business attempts to classify the business as a member of a category, such as ?restaurants,?
-?beaches,?
-or ?shopping.?
-However some categories are extremely broad and the same type of business can be described in different ways.
-One example is that a bakery may be classified as belonging to ?bakeries,?
-but also generically as ?food.?
-Conversely, many businesses sell food, but the generic category ?food?
-will not be in their list of categories and instead the business will be labelled with more specific categories like ?Italian food,?
-diminishing the usefulness of these categories.
+\section{Dataset Information}
 
-\quad The dataset provided by the Yelp organizers is fairly clean already, and the majority of the data cleaning process will be involved in handling missing data.
-It is possible to fill in these missing values with default values like ?false?, or ?none?
-but it can be dangerous to make assumptions about what businesses have and don?t have - for example, filling in a ?gluten-free?
-attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.
-
-\quad The Yelp dataset contains up to a million different attributes about businesses from 11 different cities in 4 countries.
-The majority of these attributes will be unnecessary so there will be a large effort in data reduction.
-For instance, the dataset includes photographs of the exterior and interior of some businesses (also many images of various kinds of food served by the businesses), but our questions does not require image analysis.
-There also redundant attributes such as postal codes that are unnecessary because latitude and longitude is more accurate.
-
-\subsection{Dataset}
-
-The Yelp dataset comes from the Yelp Dataset Challenge.
+\quad The Yelp dataset comes from the Yelp Dataset Challenge.
 This is an ongoing challenge and this is the ninth round, so the challenge is very well-established by this point.
@@
 
-\quad We will be considering the following from the Yelp dataset:
+\quad We considered the following from the Yelp dataset:
 
@@
 
-\section{Methods and Tools}
+\section{Main Techniques}
 
-\quad The tools and methods used for this project will no longer include a database for storing or integrating the data.
-Due to difficulties described in section 6 and relating to the unknown JSON variant used by the preparers of the Yelp dataset, storing the data in a database is not feasible at this time.
-Given that the data is already quite clean save for the format, with minimal replacement of missing values and some other changes necessary, it is possible for each group member to have a copy of the data stored locally that may be processed with our tools, with smaller subsets of data, results, and scripts shared via GitHub.
-The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?s machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical.
-We will still use various Python libraries and modules for cleaning and preprocessing, and various other tools for transformation, mining, pattern evaluation, and knowledge presentation.
-The lack of a centralized database in removes the need for a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing, but increases the necessity of group communication to reduce duplication of work and miscommunications.
-We currently use and will continue to use Slack, email, and in-person meetings to communicate and synchronize our efforts.
+\subsection{Data Cleaning and Preprocessing}
 
-\quad Numpy, Scipy, and Pandas will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation.
-Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python?s Natural Language Toolkit for sentiment analysis and other language processing.
+\quad Although the Yelp dataset was well formatted, its content was inconsistent.
+A lot of data was absent for nearly every attribute.
+Only a few attributes are guaranteed to exist for every business such as latitude and longitude.
+Businesses also contained an attributes attribute that had miscellaneous information about businesses; some examples of fields in attributes are Free Wifi, Garage Parking, etc.
+These are completely optional and are inconsistent among businesses.
+As a result of this inconsistency, we removed the attributes attribute from the dataset.
 
-\section{Evaluation}
-
-\subsection{Methodology}
-
-\quad Evaluating our predictions as to whether a business success will hinge on comparing the star rating (from one to five stars) and our predicted star rating based on the other attributes in the review, such as day of the week and month, season, or location.
-This could be done by training an ML algorithm like Kernel Ridge regression on 80\% of the data and testing it on the remaining 20\%, and then adjusting parameters to minimize the error between the actual star ratings and our predicted star ratings for the testing data.
-
-\quad Also, evaluating the sentiment analysis of the review text will likely be simple given that we will be using tools like NLTK (see following section) to perform sentiment analysis rather than creating entirely custom algorithms.
-These tools incorporate helpful metrics to analyze the error of a particular output.
-Alternatively, we can manually verify the sentiment of a (relatively small) number of reviews by reading them and comparing our human judgements of the sentiment to the output of the tools.
-For evaluating recurring problems, we will employ a similar approach, except we will group reviews by business and analyze them as a time series to and recurring issues.
-For recurring problems specifically, the most expedient method of evaluation may well be to read a small number of reviews and manually determine evaluate the performance of our tools in detecting recurrent issues.
-
-\quad For all three of the objective questions, employing association rule frameworks that use support-confidence frameworks will be helpful to find and measure strong associations, while null-invariant measures like the Kulczynski measure or the cosine measure will be helpful to determine how interesting these associations are.
-
-\section{Progress}
-
-There have been several major changes to the project tools and methods since the project proposal.
-Initially we planned on using a relational database to store, integrate, and access the data, but but we encountered many difficulties in just uploading the data.
-The inconsistencies between the user and business data made it exceedingly difficult to successfully import the data en masse into the database, so instead we decided to import the data straight into Weka.
-However, the format of the JSON file also did not allow a successful import into Weka.
-
-\quad Next we converted the JSON file into CSV format using a Python script and tried to upload it to Weka again.
-Once again Weka had issues due to the mismatch in attributes for each business (some had more attributes while others had less, likely due to the ?miscellaneous?
-category mentioned in section 3.1).
-
-\quad Instead we opted to use a different tool, RapidMiner.
-Like Weka, RapidMiner incorporates extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining.
+\quad The format of the data did not allow a successful import into Weka either in JSON or in CSV format, so instead we used RapidMiner, which incorporates extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining.
 RapidMiner is commercial software, but an educational version is available that removes certain limitations on the free version like only being able to process 10,000 rows at a time.
-Additionally, RapidMiner is much more compatible with the missing data in our dataset and automatically corrects the formatting issues we had when attempting to use Weka.
+Additionally, RapidMiner was much more compatible with the missing data in our dataset and automatically corrects some of the formatting issues we had when attempting to use Weka.
 It does not support JSON so we reused the conversion code (found on the team GitHub as convert.py) to convert the data files to CSV.
 
-\section{Results}
+\quad The remaining missing data from other attributes such as categories or address were more rare than the attributes attribute.
+Instead of removing the attribute completely, they were autocompleted with default values if certain data values were non-existent.
+For example, the categories attribute would be set a default value of null if a business lacks a category.
+These cases are rare at between 200 and 300 per attribute which is small compared to the 144,000 businesses.
 
-\subsection{Businesses}
+\quad In other cases only the subset of attributes were used to determine clusters using k-means.
+This because k-means fails to work on non-numerical values.
+So attributes such as address or name was completely avoided.
+In the case of k-means the attributes used were latitude, longitude, review count, and stars.
+These were the only numerical attributes that was provided in the businesses dataset.
 
-After importing the CSV version of the data, we tested some basic visualization tools from RapidMiner.
+\subsection{Decision Trees and K-means Clustering}
+
+\quad The decision tree was used on the Yelp dataset to breakdown classification for the category of a business.
+The decision tree was used to help categorize businesses based on the existing categories attribute for each business.
+Using the categories section as a label helped create a decision tree which split mainly on the is_open, latitude, and longitude attribute.
+These attributes yielded the highest information gain.
+Since a decision tree requires a label, the categories section became most logical to classify a business.
+As a result the cleaning of the categories attribute was important because it required no missing values to successfully label all businesses.
+
+\quad K-means clustering with k = 5 was also used to find clusters.
+In order to perform k-means it was necessary to have numeric attributes only.
+This would mean removing attributes like the address.
+In addition to this the dataset was also missing 67-166 values.
+Once again this is relatively low number so it was decided to fill in the missing values with the average of the appropriate columns.
+
+\subsection{Correlation Calculations}
+
+\quad In order to better understand the users of Yelp and draw more complete conclusions from our findings, we did a brief analysis on our user data.
+Each user has an average of the stars they give in their reviews and the number of different types of compliments they’ve received on their reviews.
+Users also have a number of fans.
+We first created a correlation matrix with RapidMiner in order to see the relationship between average stars and compliments.
+Following this a new feature was created called “total_compliments”, which was a summation of the compliments that a user has received.
+We then created a new correlation matrix to see what the overall relationship was between average stars and total compliments.
+Finally, we created another correlation matrix to examine the relationship between fans and total compliments.
+The correlation matrices were exported from RapidMiner to an image by first converting the matrix to an example data type using the ‘Converters’ extension and then writing that data to an excel file.
+The table can then be formatted in the excel file and exported to an image.
+
+\subsection{Natural Language Processing}
+
+\quad To investigate the linguistic characteristics of the review corpus, a good deal of preprocessing was necessary.
+At the beginning, all of the reviews were in one 3.46-GB file, which consistently crashed RapidMiner and necessitated the use of custom scripts focused on extracting and analyzing the information of interest.
+Since the review file was too large to open with RapidMiner, a different method of “joining” the business data and the review data was necessary.
+The script extract_businesses_many_reviews.py was used to locate businesses with more than 1000 reviews, and write the business names, IDs (a unique tag provided by Yelp), and other selected attributes like that business’s average star rating to a separate file.
+The number 1000 was chosen somewhat arbitrarily, but the following processing is the most informative for businesses with many reviews.
+There are 177 businesses with more than 1000 reviews in the dataset.
+
+\quad The script save_reviews_to_business_files.py was used to save all reviews for all businesses in the commonly-reviewed list to separate files, such that each file, named <business_id>.json, contained all the reviews for the business with that business ID.
+This step was important in building the processing pipeline, but could easily be eliminated now that the pipeline has been developed, so reviews for a certain business could be identified while the script loops through the large review file and then processed without an extra “save” step in between.
+
+\quad Once all reviews for a business were in one file, the script problems_over_time.py was used to analyze the reviews and graph the results.
+To accomplish this, each review in a <business_id>.json file with a star rating of 3 or fewer stars was tokenized, stemmed, and compared to the following list of words associated with traumatic narratives in Fig.4 on the left:
+
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{table_trauma_keywords}
+\end{figure}
+
+\quad For comparison, the list of 50 words identified by Jurafsky et al.
+(2014) is on the right (see Related Work, Narratives and Consumer Sentiment).
+Notably, none of the seven cities used in the paper were among the set of cities provided by the Yelp Dataset Challenge, so there was a guaranteed overlap of zero reviews between the two corpora.
+Also, the list of words used for problem tracking over time was stemmed, eliminating the need to include plurals and different forms like “managers” as well as “manager,” or “waiting” and “wait” as well as “waited.” The count of stemmed forms of each of the words in the above left table was normalized by the total number of reviews for that month and graphed over time.
+
+\section{Key Results}
+
+\subsection{Basic Statistics and RapidMiner}
+
+\quad After importing the CSV version of the data into RapidMiner, we tested some basic visualization tools included with the package.
 For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it is located.
@@
 For example, why does Illinois have significantly lower average stars than other states or provinces, even those that border it?
-
-\quad The most pervasive issue with our business data is that many entries are missing data, most notably the miscellaneous 'attributes' and address information.
-Since these values are managed by the account owner, they vary widely across the businesses.
-However, some of these data are redundant given the other fields, such as detailed latitude-longitude data for each business.
 
@@
 
-RapidMiner can also produce 3D scatterplots which is useful because we plan to use latitude and longitude to locate businesses, since some of the business address data is missing and cannot (and should not) be replaced with default values or other common values used to replace missing data.
+\quad RapidMiner can also produce 3D scatterplots which is useful because we used latitude and longitude in conjunction with the city and state to locate businesses, since some of the business address data is missing and cannot be filled with default values or other common values used to replace missing data.
 Figure 2 depicts a graph of longitude vs.
@@
 
-In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expect.
+\quad In this instance the visualization makes it clear that reviews are more frequent in populated cities as we expected.
 However, it also shows visually the outliers.
@@
 In this context it is easier to identify outliers and also see that businesses concentrate in certain locations, as is expected given Yelp?s declaration that the dataset focuses on businesses in 11 cities.
-Therefore we can reduce the dataset to only the cities that meet the threshold of businesses or total review counts.
+Therefore we reduced the dataset to only the cities that meet the threshold of businesses or total review counts.
 
 \subsection{Business Categories}
+
 \quad In an effort to garner more information about user behavior, we needed to augment the initial data with that from the businesses and reviews, the latter of which serves as the link between the former two.
@@
 The categories proved to be a more manageable problem as they are selected from a finite set of options.
-With numbers of selected categories ranging from two to twenty-eight out of over 1000 possible choices, we sought to reduce this high degree of dimensionality.
+With numbers of selected categories ranging from two to 28 out of over 1000 possible choices, we sought to reduce this high degree of dimensionality.
+
 \quad Sparse binary data of this kind quickly becomes unwieldy as traditional distance measures are highly prone to error.
-Such measures include K-Means, PCA analysis (for dimensionality reduction), DBSCAN, and Agglomerative Clustering.
+Such measures include k-Means, PCA analysis (for dimensionality reduction), DBSCAN, and agglomerative clustering.
 Typically, the computational cost is prohibitive (Plumpley, et.
-al., 2002) or the numerical values are highly unstable during computation.
+al.
+(2002)) or the numerical values are highly unstable during computation.
 The goal in our application is to reduce the effective categorical space and classify the businesses into a narrower subset?on the order of ~50.
@@
 Almost every form of textual representation falls into the ?bag-of-words?
-paradigm where vectors are used to represent each document, and each element corresponds to either the frequency of a given term in the dictionary (i.e.
-all words from all documents) or the binary existence of the word in the document.
+paradigm where vectors are used to represent each document, and each element corresponds to either the frequency of a given term in the dictionary (i.e., all words from all documents) or the binary existence of the word in the document.
 Since each document is highly unlikely to have an instance of all terms, the combination of these vectors results in a very large, very sparse matrix of n documents by p terms.
+
 \quad Much research has been done in recent years on Latent Semantic Analysis (or Indexing).
@@
 
-\subsection{Reviews}
+\begin{figure}[!h]
+\includegraphics[width=0.75\textwidth, center]{topics}
+\end{figure}
 
-One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time.
-However, given that we were unable to find a suitable workaround for our difficulties with Weka, we used sentiment analysis tools from the Natural Language Toolkit (NLTK) to analyze the review language instead.
-NLTK is a language-processing and analysis platform that uses Python, and a good replacement for the language processing capabilities of Weka.
+\subsection{Decision Trees}
 
-\quad Using a custom script (see GitHub\; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews.
-The remaining 5737 reviews had composite sentiment scores within -0.2 < score < 0.2, so they were excluded from further analysis since they were neutral.
-This means that the average Yelp review (at least within the analyzed subset) is positive rather than negative or neutral in sentiment, perhaps indicating that Yelp users are more inclined to write reviews of businesses with which they had a good experience.
-Only around 2.5\% of the data is currently processed by the file, because processing the entire file is prohibitively expensive.
+\quad The decision tree shows the separation of businesses by categories.
+The tree provided insight into where the categories are dominant.
+As seen in Fig.
+5 the tree breaks down into ranges of different latitude and longitude.
+The location of business has a strong correlation with the type of business.
+For instance in Fig.
+5 the tree shows that a latitude of greater than 43.784 and a longitude of less than or equal to -73.477 yielded a large amount of categories.
+This is because this coordinate is in between New York City and Montreal which are high dense cities.
+The decision tree shows the best paths to take to determine the category of a business.
+The tree relies entirely on three attributes to determine a business’s category, is_open, latitude, and longitude.
+The location of a business is important so it is logical that a more dense variety of categories appear between Montreal and New York City as opposed to Arizona where there are fewer categories.
 
-\subsection{Users}
+\quad The more interesting finding was that is_open is one of the main contributors to the category of a business.
+In retrospect the connection is clear because the failure of a business can have a domino effect.
+It is common to see a plaza or circle of businesses close down together because the location lacks popularity.
+Interestingly the tree prioritizes the is_open attribute higher than either the latitude or the longitude.
+If a business is open or closed permanently can actually determine the location and category of the business.
+One way of interpreting this is that it is very likely for a business to be a certain category if it is open or closed.
+The location is reflective of the trends and needs of that community.
+As a result many businesses of that type will either thrive or close in a certain latitude and longitude.
+Using the Montreal example before, the Canadian city has influence from New York City below it, the French, and of course Canadian culture itself.
+This blend of culture is well reflected in the culinary culture of Montreal.
+The tree in Fig.
+5 portrays this area having categories of Food Delis, Comfort Food, Restaurants, Creperies, and Desserts.
+If a business is open and falls within those coordinates then it is likely that the business serves food of some sort.
 
-The users dataset essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem.
-The figure below shows a plot of the user data.
-We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews.
+\quad By contrast, the closed businesses tend to be more specialized and have an exact purpose within a certain community.
+Their locations are in less populated cities.
+For instance in Fig.
+6, businesses that are categorized as Heating and Air Conditioning is in a less populated area.
+The color key below the node indicates the ratio of competing businesses in that same area.
+In the case of Heating and Air Conditioning it is the monopoly in its area.
+The business specializes in Heating and Air Conditioning where only a single type of this business is necessary in a smaller community.
+It is also follows that these locations are more obscure because they exist in areas where changes in weather are more extreme requiring the aid of this business.
+
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{decision_tree_open}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{decision_tree_closed}
+\end{figure}
+
+\quad The use of k-means returned results that were surprisingly consistent in star reviews (Fig.
+7).
+Many of the businesses are in highly populated cities.
+As a result the k-means struggled to separate these businesses by location.
+However, k-means reveals the significant difference in average review counts by location.
+For instance the average review count for cluster 0 is ~207 reviews compared to cluster 1 which has ~18 reviews on average.
+However the average star rating is roughly similar around 3.7 stars.
+This pattern is seen in the other three clusters where differences in review counts still result in similar average star ratings.
+Across all clusters the stars are all near 3.7 stars average reviews.
+In cluster 2 the average star rating is ~3.5 which varies more significantly from the other clusters.
+The responding latitude and longitude is in the Massachusetts and New Hampshire area.
+The east coast has significantly less average review, but also lower star ratings as seen before.
+This may indicate more strict culinary culture on the east coast compared to the west with a relaxed comfort food culture.
+Yelp is a Silicon Valley company so its adoption and usage is likely more significant in western states.
+
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{k_means_centroids}
+\end{figure}
+
+\subsection{User Behavior}
+
+\quad The users dataset essentially provides a window into users' status, connections, and interactions within the Yelp ecosystem.
+The figure below shows a plot of the user density versus the average star rating.
+We see that the density grows around 3.25-4.25 stars with some traces of a bell curve regarding the number of reviews the user has written.
 
@@
 
-\quad One of the most significant yet hardest to process attributes is the 'friends' list.
-This attribute provides an opportunity to model our data as a graph.
-Such a network could provide very relevant clustering through sparseness measures and cutting in addition to accounting for user/business influence or finding deeper trends according to network connectivity.
-However, given the magnitude of the dataset, modeling the ?friends?
-of each user as a complete graph will be computationally difficult.
-To combat this, we could limit our model to the first n people the graph builder encounters only, since connections between groups of people are unpredictable.
+\quad Our correlation matrices for the user data showed some insight into the social behavior on Yelp.
+Our first matrix included each complement type and the average stars a user gave in their reviews (Fig.
+8).
+This matrix showed that there was not a correlation between the number of stars a user would give in their reviews and the compliments they receive.
+This seems to indicate that users would not be motivated to give a business a better or worse review in order to gain favorability in their profile.
 
-\quad
-\section{Next Steps}
+\quad The second matrix (Fig.
+9) includes the average stars feature and the total compliments feature.
+Again, there does not seem to be a relationship between star ratings given and the compliments received.
 
-\quad It is still necessary to reduce dimensionality.
-We've currently been working on the whole dataset and simply ignoring all attributes that fall in the category 'other.' This is cumbersome and it will become increasingly detrimental as we use more advanced techniques like clustering.
-The most prioritized attribute is the attributes portion.
-This attribute has the most issues because it is a miscellaneous catch-all where any attribute that didn?t fit with the others is thrown in.
-The issue is that it totally lacks consistency in the number and types of miscellaneous attributes included for each business.
-One business may have a Free Wifi attribute while others may not.
-After looking through the data it was clear that there wasn?t one attribute in this misc.
-category that was reliably present for all businesses.
-Although we have moved on to work with the other attributes, this one still needs work to decide if it should be omitted completely or if default values should be set instead.
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{correlation_matrix}
+\end{figure}
 
-\quad In addition, we need to combine the businesses and reviews dataset.
-This becomes extremely useful when examining the sentiment of a review and finding a correlation with the attributes of a business.
-This helps us build a model to potentially predict if a business receives negative or positive reviews based on its attributes.
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{star_and_compliment_matrix}
+\end{figure}
 
-\quad As mentioned before, we will be interested in clustering businesses together as to whether they receive positive or negative reviews based on their attributes.
-This can be difficult with the high number of dimensions of these datasets.
-We will try CLIQUE and PROCLUS for subspace clustering.
-PROCLUS will be particularly useful because the subspace for one business to qualify as ?positively reviewed?
-could be different than the subspace for another business.
-This allows us to process this data faster, but also consider the complexities of various subspaces.
+\subsection{Tracking Problems}
 
-\quad The user and review datasets are the largest we have.
-We would like to use these to find useful information about the people and the reviews they give.
-Because Yelp is a platform on which anybody can create an account and write a review, it is possible that not everybody is honest and fair in their reviews.
-Furthermore, different people may assign scores for different reasons.
+\quad To answer the question of what types of problems were repeatedly present in a business, we built a utility that analyzes the language found in reviews.
+As mentioned in the Related Work section, Jurafsky et al.
+(2014) used a log-odds-ratio informative Dirichlet prior to determine the 50 words most statistically overrepresented in Yelp reviews with one star.
+We then used the stemmed forms of most of these words (see Fig.
+4 in Main Techniques, Natural Language Processing) to determine the “issue distribution” for a particular business during a particular month, two examples of which are below as Fig.
+10:
 
-\quad Some useful attributes in the user data alone are the number of reviews that an individual has made, the length of time a user has been active on Yelp, and whether their reviews tend to be tagged as useful by other users.
+\begin{figure}[h]
+\includegraphics[width=0.75\textwidth, center]{nltk_tracking}
+\end{figure}
 
-\quad One additional consideration is processing power.
-Our chosen tool seems to be very computationally intensive, so we will very likely need to use sampled data to test hypotheses then turn to python scripts to actually run over the entirety of our training, testing, and validations sets.
-Furthermore, our data is currently only linked through review, user, and business id pointers.
-While there?s a great deal to be learned from each collection individually, we will need to connect our disparate sets by these terms.
+\quad The high variability in weight seen in from 2008-2010 for Earl of Sandwich is due to a higher prevalence of negative words during every other month, which was a common occurrence in many of the analyzed restaurants.
+Restaurants were the only type of business from the review data that this analysis was applied to, because the “training set” of words provided by Jurafsky et al.
+(2014) were obtained from an analysis of restaurants.
+The spikes every other month could be the result of customers with high hopes visiting and coming away disappointed, leaving waves of more positive and more negative reviews.
+Additionally, it is worth noting that the average star rating for Bachi Burger within the dataset is 4.0, and the corresponding value for Earl of Sandwich is 4.5.
+Although both businesses had greater than 2700 reviews in the dataset, the number with 3 stars or fewer was substantially fewer, limiting the effectiveness of the analysis and providing another mechanism for the spikes seen in both graphs.
 
-\quad Currently, the review dataset is too large to process in detail.
-In the future, we will continue taking subset samples of a more reasonable size and processing those instead, as well as increasing the complexity of our analysis.
-Combining the review dataset with the business dataset (using the 'business\_id' field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews.
-We hope that additional advanced sentiment analysis will be useful in determining underlying factors that influence a business's success, as well as in identifying recurrent problems.
+\section{Applications}
 
-\quad Once we have our sentiment analysis we can combine this data with our user data and business data to find patterns in the sentiment of reviews.
-We would like to compare the sentiment of reviews to the area they are coming from, the businesses they apply to, and whether reviews of a given sentiment are considered beneficial by other users.
-Using this information we can then narrow down the factors that are most influential to the score of a business in order to draw our final conclusions.
+\quad The decision tree provides insight into the separation of businesses by their category.
+This can help classify the category of a new business or new areas.
+For instance, given a new business located at x latitude and y longitude, Yelp can make predictions based on surrounding categories.
+The tree also reveals the significance of a business being open or permanently closed, and potential declines in certain communities of businesses.
+Communities that have closing businesses are likely to have surrounding businesses close as well.
+In terms of businesses, this can help entrepreneurs choose areas that aren’t in decline when starting a new business.
+
+\quad Latitude and longitude are also significant to the category of a business.
+For businesses this is important because certain categories become abundant in certain areas.
+One example is that in both New York City and Montreal there are a significant number of creperies because of New York City’s large population density and Montreal’s French influences.
+New businesses should consider other locations where creperies aren’t as popular.
+This can provide opportunities for new or existing businesses to venture into new areas.
+It can also help reveal the niche in certain communities such as in Fig.
+4 where certain areas have relatively high concentrations of fitness-related businesses and even martial arts.
+New businesses can add to concentrations of businesses with an extra twist or bring quality to a category.
+This quality can be better determined with the k-means clusters.
+
+\quad The clusters determined by k-means portray where Yelp is more impactful to a business.
+West coast areas receive more reviews on average.
+This can be helpful for businesses to measure their performance.
+Businesses that want to receive more feedback may consider the clusters that have higher review counts.
+It also reveals a more active community of Yelp which could be better business.
+There may be more competition in this clusters, but success can result in exponential growth in an active Yelp community.
+
+\quad As mentioned in the introduction, the natural language processing performed on the review text to identify negative sentiment and recurring problems can be leveraged by managers and other staff members to improve their ratings, thereby improving revenue.
 
@@
 
-%\section{Typical references in new ACM Reference Format}
-%A paginated journal article \cite{Abril07}, an enumerated
-%journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
-%a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
-%\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
-%followed by the same example, however we only output the series if the volume number is given
-%\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
-%a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
-%in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
-%an article in a proceedings (of a conference, symposium, workshop for example)
-%(paginated proceedings article) \cite{Andler79}, a proceedings article
-%with all possible elements \cite{Smith10}, an example of an enumerated
-%proceedings article \cite{VanGundy07},
-%an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
-%a master's thesis: \cite{anisi03}, an online document / world wide web
-%resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
-%and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
-%work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
-%\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
-%'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
-%Boris / Barbara Beeton: multi-volume works as books
-%\cite{MR781536} and \cite{MR781537}.
-%
-%A couple of citations with DOIs: \cite{2004:ITE:1009386.1010128,
-%  Kirschmer:2010:AEI:1958016.1958018}.
+\section{Typical references in new ACM Reference Format}
+
+A paginated journal article \citeA{Abril07}, an enumerated
+journal article \citeA{Cohen07}, a reference to an entire issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
+\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
+followed by the same example, however we only output the series if the volume number is given
+\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
+a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
+in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
+an article in a proceedings (of a conference, symposium, workshop for example)
+(paginated proceedings article) \cite{Andler79}, a proceedings article
+with all possible elements \cite{Smith10}, an example of an enumerated
+proceedings article \cite{VanGundy07},
+an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
 
 % Bibliography
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{ACM-Reference-Format}{}
 \bibliography{sample-bibliography}

### 2017-05-02 paper/paper.tex
final compilation

@@
 Decision trees and k-means clustering were employed to group businesses by attributes with the highest information gain or shortest distance in attribute space, respectively.
-Furthermore, sentiment analysis was used on the language present in the reviews and compared with the star reviews to provide more information on a business’s chances of success.
+Furthermore, sentiment analysis was used on the language present in the reviews and compared with the star reviews to provide more information on a business's chances of success.
 Information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.
@@
 This type of information will be less useful to businesses, but it could prove very helpful for customers or advertisers to know that businesses in a certain area tend to be reviewed much higher or lower than the average business in cities of a comparable size.
-Also, to take an example, businesses reading their reviews on Yelp would be aware that the reviews in the area tend to be more negative than is average for the state, which in turn would prevent them from trying to fix problems that don’t exist.
+Also, to take an example, businesses reading their reviews on Yelp would be aware that the reviews in the area tend to be more negative than is average for the state, which in turn would prevent them from trying to fix problems that don't exist.
 
@@
 \quad A previous winner, Hood et al.
-(2013) [2], published a paper on assessing a business’s current rating/review state and future expectations with respect to review numbers on Yelp.
+(2013) [2], published a paper on assessing a business's current rating/review state and future expectations with respect to review numbers on Yelp.
 The bulk of their work centered around the creation of additional features for the data (e.g., the number of similar businesses within 1 km, features of subsets of reviews such as average rating and number of unique users, etc.).
@@
 Our approach was similar in the application of feature reduction to reduce data dimensionality, and the use of clustering to not only find cities with many businesses, but also group similar businesses together.
-However, save for the total_compliments feature created to help analyze the user data, we did not add additional features to the dataset like the number of nearby similar businesses.
+However, save for the total\_compliments feature created to help analyze the user data, we did not add additional features to the dataset like the number of nearby similar businesses.
 
@@
 (2014) [6] determined that negative reviews are likely to incorporate words and other features previously associated with trauma.
-Specifically, the incidence of words and phrases with a negative emotional slant, words relating to actions performed by third-person actors like waiters or management, and first-person plural pronouns like “we” and “us” is higher for negative reviews than positive.
+Specifically, the incidence of words and phrases with a negative emotional slant, words relating to actions performed by third-person actors like waiters or management, and first-person plural pronouns like 'we' and 'us' is higher for negative reviews than positive.
 Meanwhile, positive reviews were characterized by language associated with addiction and, in the case of more expensive restaurants, sensory pleasure.
 
-\quad To determine precisely which words were statistically overrepresented in the category of negative reviews, in this case taken to be reviews with one star, the authors applied a “log-odds-ratio informative Dirichlet prior” (originally published in Monroe et al.
+\quad To determine precisely which words were statistically overrepresented in the category of negative reviews, in this case taken to be reviews with one star, the authors applied a 'log-odds-ratio informative Dirichlet prior' (originally published in Monroe et al.
 (2008) [7]) to compare the frequency of each word in the full corpus to its frequency in the category of one-star reviews.
 The authors then sorted all the words by their log-odds association score to determine the 50 words that were most associated with one-star reviews.
-Rather than repeat their work, our group used a number of these words as given in the paper to determine the “issue distribution” of a particular restaurant over time (see Key Results, Tracking Problems).
+Rather than repeat their work, our group used a number of these words as given in the paper to determine the 'issue distribution' of a particular restaurant over time (see Key Results, Tracking Problems).
 
@@
 	\begin{verbatim}
-	"review_id":(encrypted review id),
-	"user_id":(encrypted user id),
-	"business_id":(encrypted business id)"
+	"review\_id":(encrypted review id),
+	"user\_id":(encrypted user id),
+	"business\_id":(encrypted business id)"
 	"stars":(star rating, rounded to half-stars),
@@
 	\begin{verbatim}
-	"business_id":(encrypted business id),
+	"business\_id":(encrypted business id),
 	"name":"business name",
@@
 	"stars":(star rating, rounded to half-stars),
-	"review_count":(number of reviews),
-	"is_open":0/1 (closed/open),
+	"review\_count":(number of reviews),
+	"is\_open":0/1 (closed/open),
 	"attributes":[(localized attribute tags)],
@@
 	\begin{verbatim}
-	"user_id":(encrypted user id),
+	"user\_id":(encrypted user id),
 	"name":"first name",
-	"review_count":(number of reviews),
-	"yelping_since": (date, formatted: YYYY-MM-DD),
+	"review\_count":(number of reviews),
+	"yelping\_since": (date, formatted: YYYY-MM-DD),
 	"friends":[(encrypted user ids)],
@@
 	"elite":["an array of years the user was elite"],
-	"average_stars":floating point average like 4.31,
-	"compliment_type": (compliment count),
+	"average\_stars":floating point average like 4.31,
+	"compliment\_type": (compliment count),
 		...(same for each complement type)...
@@
 The decision tree was used to help categorize businesses based on the existing categories attribute for each business.
-Using the categories section as a label helped create a decision tree which split mainly on the is_open, latitude, and longitude attribute.
+Using the categories section as a label helped create a decision tree which split mainly on the is\_open, latitude, and longitude attribute.
 These attributes yielded the highest information gain.
@@
 \quad In order to better understand the users of Yelp and draw more complete conclusions from our findings, we did a brief analysis on our user data.
-Each user has an average of the stars they give in their reviews and the number of different types of compliments they’ve received on their reviews.
+Each user has an average of the stars they give in their reviews and the number of different types of compliments they've received on their reviews.
 Users also have a number of fans.
 We first created a correlation matrix with RapidMiner in order to see the relationship between average stars and compliments.
-Following this a new feature was created called “total_compliments”, which was a summation of the compliments that a user has received.
+Following this a new feature was created called 'total\_compliments', which was a summation of the compliments that a user has received.
 We then created a new correlation matrix to see what the overall relationship was between average stars and total compliments.
 Finally, we created another correlation matrix to examine the relationship between fans and total compliments.
-The correlation matrices were exported from RapidMiner to an image by first converting the matrix to an example data type using the ‘Converters’ extension and then writing that data to an excel file.
+The correlation matrices were exported from RapidMiner to an image by first converting the matrix to an example data type using the '€˜Converters' extension and then writing that data to an excel file.
 The table can then be formatted in the excel file and exported to an image.
@@
 At the beginning, all of the reviews were in one 3.46-GB file, which consistently crashed RapidMiner and necessitated the use of custom scripts focused on extracting and analyzing the information of interest.
-Since the review file was too large to open with RapidMiner, a different method of “joining” the business data and the review data was necessary.
-The script extract_businesses_many_reviews.py was used to locate businesses with more than 1000 reviews, and write the business names, IDs (a unique tag provided by Yelp), and other selected attributes like that business’s average star rating to a separate file.
+Since the review file was too large to open with RapidMiner, a different method of 'joining' the business data and the review data was necessary.
+The script extract\_businesses\_many\_reviews.py was used to locate businesses with more than 1000 reviews, and write the business names, IDs (a unique tag provided by Yelp), and other selected attributes like that business's average star rating to a separate file.
 The number 1000 was chosen somewhat arbitrarily, but the following processing is the most informative for businesses with many reviews.
@@
 
-\quad The script save_reviews_to_business_files.py was used to save all reviews for all businesses in the commonly-reviewed list to separate files, such that each file, named <business_id>.json, contained all the reviews for the business with that business ID.
-This step was important in building the processing pipeline, but could easily be eliminated now that the pipeline has been developed, so reviews for a certain business could be identified while the script loops through the large review file and then processed without an extra “save” step in between.
+\quad The script save\_reviews\_to\_business\_files.py was used to save all reviews for all businesses in the commonly-reviewed list to separate files, such that each file, named <business\_id>.json, contained all the reviews for the business with that business ID.
+This step was important in building the processing pipeline, but could easily be eliminated now that the pipeline has been developed, so reviews for a certain business could be identified while the script loops through the large review file and then processed without an extra 'save' step in between.
 
-\quad Once all reviews for a business were in one file, the script problems_over_time.py was used to analyze the reviews and graph the results.
-To accomplish this, each review in a <business_id>.json file with a star rating of 3 or fewer stars was tokenized, stemmed, and compared to the following list of words associated with traumatic narratives in Fig.
+\quad Once all reviews for a business were in one file, the script problems\_over\_time.py was used to analyze the reviews and graph the results.
+To accomplish this, each review in a <business\_id>.json file with a star rating of 3 or fewer stars was tokenized, stemmed, and compared to the following list of words associated with traumatic narratives in Fig.
 1 on the left:
@@
 Notably, none of the seven cities used in the paper were among the set of cities provided by the Yelp Dataset Challenge, so there was a guaranteed overlap of zero reviews between the two corpora.
-Also, the list of words used for problem tracking over time was stemmed, eliminating the need to include plurals and different forms like “managers” as well as “manager,” or “waiting” and “wait” as well as “waited.” The count of stemmed forms of each of the words in the above left table was normalized by the total number of reviews for that month and graphed over time.
+Also, the list of words used for problem tracking over time was stemmed, eliminating the need to include plurals and different forms like 'managers' as well as 'manager,' or 'waiting' and 'wait' as well as 'waited.' The count of stemmed forms of each of the words in the above left table was normalized by the total number of reviews for that month and graphed over time.
 
@@
 There are various businesses that have only a few reviews.
-In this context it is easier to identify outliers and also see that businesses concentrate in certain locations, as is expected given Yelp?s declaration that the dataset focuses on businesses in 11 cities.
+In this context it is easier to identify outliers and also see that businesses concentrate in certain locations, as is expected given Yelp's declaration that the dataset focuses on businesses in 11 cities.
 Therefore we reduced the dataset to only the cities that meet the threshold of businesses or total review counts.
@@
 In doing so, we attempted to profile the most frequent business reviewed by each user.
-The two troublesome business features were the user-defined ?tags?
-and ?categories?.
+The two troublesome business features were the user-defined 'tags' and 'categories'.
 The categories proved to be a more manageable problem as they are selected from a finite set of options.
@@
 Such measures include k-means, PCA analysis (for dimensionality reduction), DBSCAN, and agglomerative clustering.
-Typically, the computational cost is prohibitive (Plumpley, et.
-al.
-(2002)) or the numerical values are highly unstable during computation.
+Typically, the computational cost is prohibitive (Plumley, (2002)) or the numerical values are highly unstable during computation.
 The goal in our application is to reduce the effective categorical space and classify the businesses into a narrower subset?on the order of ~50.
@@
 The decision tree shows the best paths to take to determine the category of a business.
-The tree relies entirely on three attributes to determine a business’s category, is_open, latitude, and longitude.
+The tree relies entirely on three attributes to determine a business's category, is\_open, latitude, and longitude.
 The location of a business is important so it is logical that a more dense variety of categories appear between Montreal and New York City as opposed to Arizona where there are fewer categories.
 
-\quad The more interesting finding was that is_open is one of the main contributors to the category of a business.
+\quad The more interesting finding was that is\_open is one of the main contributors to the category of a business.
 In retrospect the connection is clear because the failure of a business can have a domino effect.
 It is common to see a plaza or circle of businesses close down together because the location lacks popularity.
-Interestingly the tree prioritizes the is_open attribute higher than either the latitude or the longitude.
+Interestingly the tree prioritizes the is\_open attribute higher than either the latitude or the longitude.
 If a business is open or closed permanently can actually determine the location and category of the business.
@@
 We then used the stemmed forms of most of these words (see Fig.
-1 in Main Techniques, Natural Language Processing) to determine the “issue distribution” for a particular business during a particular month, two examples of which are below as Fig.
+1 in Main Techniques, Natural Language Processing) to determine the 'issue distribution' for a particular business during a particular month, two examples of which are below as Fig.
 10:
@@
 \quad The high variability in weight seen in from 2008-2010 for Earl of Sandwich is due to a higher prevalence of negative words during every other month, which was a common occurrence in many of the analyzed restaurants.
-Restaurants were the only type of business from the review data that this analysis was applied to, because the “training set” of words provided by Jurafsky et al.
+Restaurants were the only type of business from the review data that this analysis was applied to, because the 'training set' of words provided by Jurafsky et al.
 (2014) were obtained from an analysis of restaurants.
@@
 Communities that have closing businesses are likely to have surrounding businesses close as well.
-In terms of businesses, this can help entrepreneurs choose areas that aren’t in decline when starting a new business.
+In terms of businesses, this can help entrepreneurs choose areas that aren't in decline when starting a new business.
 
@@
 For businesses this is important because certain categories become abundant in certain areas.
-One example is that in both New York City and Montreal there are a significant number of creperies because of New York City’s large population density and Montreal’s French influences.
-New businesses should consider other locations where creperies aren’t as popular.
+One example is that in both New York City and Montreal there are a significant number of creperies because of New York City's large population density and Montreal's French influences.
+New businesses should consider other locations where creperies aren't as popular.
 This can provide opportunities for new or existing businesses to venture into new areas.
@@
 \begin{enumerate}
-	\item Yelp Dataset Challenge, Round 9: https://www.yelp.com/dataset_challenge
+	\item Yelp Dataset Challenge, Round 9: https://www.yelp.com/dataset\_challenge
 
@@
 in The Yelp Challenge, (Carnegie Mellon University, Pittsburgh, PA, 2013), Yelp.com.
-Retrieved March 3, 2017, from Yelp.com: https://www.yelp.com/html/pdf/YelpDatasetChallengeWinner_InferringFuture.pdf.
+Retrieved March 3, 2017, from Yelp.com: https://www.yelp.com/html/pdf/YelpDatasetChallengeWinner\_InferringFuture.pdf.
 
@@
 in iConference, (Berlin, Germany, 2014), Yelp.com.
-Retrieved March 3, 2017, from Yelp.com: https://www.yelp.com/html/pdf/YelpDatasetChallengeWinner_ImprovingRestaurants.pdf.
+Retrieved March 3, 2017, from Yelp.com: \\https://www.yelp.com/html/pdf/YelpDatasetChallengeWinner\_ImprovingRestaurants.pdf.
 
@@
 Submitted to American Economic Journal: Applied Economics.
-Retrieved March 3, 2017, from Harvard Business School: http://www.hbs.edu/faculty/Publication%20Files/12-016_a7e4a5a2-03f9-490d-b093-8f951238dba2.pdf.
+Retrieved March 3, 2017, from Harvard Business School: http://www.hbs.edu/faculty/Publication\%20Files/12-016\_a7e4a5a2-03f9-490d-b093-8f951238dba2.pdf.
 
@@
 
-    \item Monroe, B., Colaresi, M., and Quinn, K. Fightin’ words: Lexical feature selection and evaluation for identifying the content of political conflict. Political Analysis, 16 (4). 372-403.
+    \item Monroe, B., Colaresi, M., and Quinn, K. Fightin' words: Lexical feature selection and evaluation for identifying the content of political conflict. Political Analysis, 16 (4). 372-403.
+
+    \item Plumbley, M.
+Clustering of Sparse Binary Data using Minimum Description Length Approach, (University of London, London, UK, 2002), Retrieved April 26, 2017 from ResearchGate: \\https://www.researchgate.net/publication/2539415\_Clustering\_of\_Sparse\_Binary\_Data\_\\using\_a\_Minimum\_Description\_Length\_Approach
 

### 2017-05-02 paper/paper.tex
final paper tweaks

@@
 Such measures include k-means, PCA analysis (for dimensionality reduction), DBSCAN, and agglomerative clustering.
-Typically, the computational cost is prohibitive (Plumley, (2002)) or the numerical values are highly unstable during computation.
-The goal in our application is to reduce the effective categorical space and classify the businesses into a narrower subset?on the order of ~50.
+Typically, the computational cost is prohibitive [8] or the numerical values are highly unstable during computation.
+The goal in our application is to reduce the effective categorical space and classify the businesses into a narrower subset\--on the order of 50.
 Not surprisingly, this scenario is almost identical to that encountered when processing and analyzing large corpuses of text.

### 2017-05-02 paper/paper.tex
last one

@@
 Not surprisingly, this scenario is almost identical to that encountered when processing and analyzing large corpuses of text.
-Almost every form of textual representation falls into the ?bag-of-words?
-paradigm where vectors are used to represent each document, and each element corresponds to either the frequency of a given term in the dictionary (i.e., all words from all documents) or the binary existence of the word in the document.
+Almost every form of textual representation falls into the 'bag-of-words' paradigm where vectors are used to represent each document, and each element corresponds to either the frequency of a given term in the dictionary (i.e., all words from all documents) or the binary existence of the word in the document.
 Since each document is highly unlikely to have an instance of all terms, the combination of these vectors results in a very large, very sparse matrix of n documents by p terms.
@@
 \quad Much research has been done in recent years on Latent Semantic Analysis (or Indexing).
-This utilizes the fact that a given matrix?s Singular Value Decomposition provides us with a large degree of useful data concerning the nature of the matrix, namely which vectors and rows carry more weight as well as those with less impact.
+This utilizes the fact that a given matrix's Singular Value Decomposition provides us with a large degree of useful data concerning the nature of the matrix, namely which vectors and rows carry more weight as well as those with less impact.
 Furthermore, it provides a model through which new data can be projected onto the reduced subspace.
